---
title: "01 Trial Data Curation"
output: html_document
date: "2024-02-07"
---


# About
mvGWAS project
Review and QC of cassava breeding trial data
-   Inputs: 
        - Field trial data from Casssavabase, HarvestPlus accessions only: `./data/phenotypes_yellowGWAS_2023-10-02CBdownload.csv `
        - Field trial data from Casssavabase, including CASS accessions: `./data/2024-02-01T130213phenotype_download.csv`
-   Outputs: BLUPs and Broad-sense Heritability estimates from cleaned field trial data (outliers removed):
        - BLUPs: `./data/alltrait_drgBLUPs_HP.RDS` and `./data/alltrait_drgBLUPs_CASS.RDS`
        - H2: `./data/alltrait_H2_HP.RDS` and `./data/alltrait_H2_CASS.RDS`
        - Trait descriptions table
        - Trial descriptions table
        - Trait heritability table


Phenotypic data QC is based on Marnin Wolfe's [NextGen Cassava GS pipeline](https://github.com/wolfemd/IITA_2021GS/) using tools from the `genomicMateSelectR` package.

# Setup
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse); library(magrittr); library(genomicMateSelectR); library(MASS); library(asreml); library(scales)
#devtools::install_github("wolfemd/genomicMateSelectR", ref = 'master') 

project_path <- "/workdir/ssv42/mvGWAS/"
setwd(project_path)
set.seed(14850)
```


# Cassavabase download information

1.  [Cassavabase search wizard](https://www.cassavabase.org/breeders/search) was used to make lists. I downloaded phenotypes and meta-data for trials in the list using the **Manage** --\> **Download** page [here](https://www.cassavabase.org/breeders/download).
2.  Trial data including CASS accessions: Selected trials which contained at least four of the genotyped accessions; locations IB, IK, UBI, UMU, MOK, or ILO; Years 2013-2023. Those trials were put into list 'yellowGWAS23_trials_updated' (473 trials). Download saved as `./data/trial_data/2024-02-01T130213phenotype_download.csv`.
3.  Trial data with only the HarvestPlus accessions: Selected trials with names containing the keywords `yrt`, `highTC`, `TC`, or `mixed` and/or accessions in the list of genotyped accessions `yellow_GWAS_23`. Those trials were put into list `yellow_GWAS_23_trials` (# trials). Download saved as `./data/phenotypes_yellowGWAS_2023-10-02CBdownload.csv`.

# Data curation

Read in data and select analysis options
```{r}
#Select dataset
dataset <- "CASS" #"HP" #or
print(paste0("working with ", dataset, " data"))

#Select filtering method
filter_trial_byH2 <- FALSE

#Select trait
TCtrait <- "TCICHK"


if(dataset=="CASS"){
  rawtrialdata <- read.csv("./data/2024-02-01T130213phenotype_download.csv")
  samplenames <- read.table("./data/genotypenames_CASS.txt")[,1]
  
} else if(dataset=="HP"){
  rawtrialdata <- read.csv("./data/phenotypes_yellowGWAS_2023-10-02CBdownload.csv")
  samplenames <- na.exclude(c(read.csv("./data/platecoordinates_HP.csv")[,"UniqueSampleID"], "IITA-TMS-IBA070593")[-c(357:384)]) #drop ITH names
}
```

Select traits of interest:
```{r, warning=FALSE, message=FALSE}
#dataset specific traitnames
if(dataset=="CASS"){
  lchromo <- "L.chromometer.of.fresh.root.CO_334.0002065"
  achromo <- "A.chromometer.value.of.fresh.root.CO_334.0002066"
  bchromo <- "B.chromometer.value.of.fresh.root.CO_334.0002064"
} else if(dataset=="HP"){
  lchromo <- "L.chromometer.value.CO_334.0002065"
  achromo <- "a.chromometer.value.CO_334.0002066"
  bchromo <- "b.chromometer.value.CO_334.0002064"
}


#define trait abbreviations
traitabbrevs <- tribble(~TraitAbbrev,~TraitName,
        "TCICHK",     'total.carotenoid.by.iCheck.method.CO_334.0000162',
        "TCHART",     'total.carotenoid.by.chart.1.8.CO_334.0000161',
        "DM",         'dry.matter.content.percentage.CO_334.0000092', 
        "DYLD",       'dry.yield.CO_334.0000014',
        "STRCH",      'starch.content.percentage.CO_334.0000071',
        "RTWT",       'fresh.storage.root.weight.per.plot.CO_334.0000012',
        "DMSPGRAV",   'dry.matter.content.by.specific.gravity.method.CO_334.0000160',
        "NOSTK",      'number.of.planted.stakes.per.plot.counting.CO_334.0000159',
        "POUND",      'poundability.assessment.0.4.CO_334.0000074',
        "PLARCH",     'plant.architecture.visual.rating.1.5.CO_334.0000099',
        "TOPYLD",     'top.yield.CO_334.0000017',
        "CGMS1",      'cassava.green.mite.severity.first.evaluation.CO_334.0000189',
        "CGMS2",      'cassava.green.mite.severity.second.evaluation.CO_334.0000190',
        "LODGE",      'proportion.lodged.plants.in.percentage.CO_334.0000094',
        "BRNHT",      'first.apical.branch.height.measurement.in.cm.CO_334.0000106',
        "BRLVLS",     'branching.level.counting.CO_334.0000079',
        "SPRTPERC",   'sprouting.proportion.CO_334.0000008',
        "HI",         'harvest.index.variable.CO_334.0000015',
        "SHTWT",      'fresh.shoot.weight.measurement.in.kg.per.plot.CO_334.0000016',
        "RTNRT",      'rotted.storage.root.counting.CO_334.0000084',
        "FRYLD",      'fresh.root.yield.CO_334.0000013',
        "RTNO",       'root.number.per.plot.counting.CO_334.0000011',
        "NOHAV",      'plant.stands.harvested.counting.CO_334.0000010',
        "CMD1S",      "cassava.mosaic.disease.severity.1.month.evaluation.CO_334.0000191",
        "CMD3S",      "cassava.mosaic.disease.severity.3.month.evaluation.CO_334.0000192",
        "CMD3I",      'cassava.mosaic.disease.incidence.3.month.evaluation.CO_334.0000196',
        "CMD6S",      "cassava.mosaic.disease.severity.6.month.evaluation.CO_334.0000194",
        "CMD9S",      "cassava.mosaic.disease.severity.9.month.evaluation.CO_334.0000193",
        "CBB3I",      'cassava.bacterial.blight.incidence.3.month.evaluation.CO_334.0000178',
        "CBB3S",      'cassava.bacterial.blight.severity.3.month.evaluation.CO_334.0000175',
        "CAD6I",      'cassava.anthractnose.disease.incidence.in.6.month.CO_334.0000181',
        "CAD6S",      'cassava.anthractnose.disease.severity.in.6.month.CO_334.0000184',
        "VIGOR",      'initial.vigor.assessment.1.7.CO_334.0000009',
        "LCHROMO",    lchromo,
        "ACHROMO",    achromo,
        "BCHROMO",    bchromo) 


#print traitnames
#traitabbrevs %>% rmarkdown::paged_table()


#save traitnames
traitnames <- as.data.frame(traitabbrevs)$TraitAbbrev
saveRDS(traitnames, file=paste0(project_path, "data/alltraitnames_", dataset, ".RDS"))

```


Clean & QC data
```{r, warning=FALSE, message=FALSE}
#define selection stage 
rawtrialdata_staged <- rawtrialdata %>%

  #Extract selection stage from trial name
  mutate(selStage = case_when(grepl("pyt", studyName, ignore.case=T) ~ "PYT",
                              grepl("ayt", studyName, ignore.case=T) ~ "AYT",
                              grepl("uyt", studyName, ignore.case=T) ~ "UYT",
                              grepl("NCRP", studyName, ignore.case=T) ~ "UYT", 
                              #NCRP = nationally coordinated research program trial
                              grepl("cet", studyName, ignore.case=T) ~ "CET",
                              grepl("\\.CE\\.", studyName, ignore.case=T) ~ "CET",
                              grepl("clonaleva", studyName, ignore.case=T) ~ "CET",
                              grepl("22.GS.C6.STUDENT2.CE.500.IK", studyName, ignore.case=T) ~ "CET",
                              grepl("22.GS.C7.CE.1360.IK", studyName, ignore.case=T) ~ "CET",
                              grepl("geneticgain", studyName, ignore.case=T) ~ "GG",
                              grepl("stock", studyName, ignore.case=T) ~ "MLT",
                              grepl("Germ", studyName, ignore.case=T) ~ "MLT",
                              grepl("conservation", studyName, ignore.case=T) ~ "MLT",
                              T ~ "NA") %>% as.factor())


# Run function `renameAndSelectCols()` to rename columns and remove everything unnecessary
# QC Trait values
trialdatall <- renameAndSelectCols(traitabbrevs, indata=rawtrialdata_staged, 
                 customColsToKeep = c("TrialType","observationUnitName","selStage")) %>%
  as.data.frame() %>% 
  filter(!is.na(germplasmName)) %>% #remove trials without germplasm names
  mutate(CMD1S=ifelse(CMD1S<1 | CMD1S>5,NA,CMD1S), #remove any nonsensical values for ordinal traits
         CMD3S=ifelse(CMD3S<1 | CMD3S>5,NA,CMD3S),
         CMD6S=ifelse(CMD6S<1 | CMD6S>5,NA,CMD6S), 
         CMD9S=ifelse(CMD9S<1 | CMD9S>5,NA,CMD9S),
         CGMS1=ifelse(CGMS1<1 | CGMS1>5,NA,CGMS1),
         CGMS2=ifelse(CGMS2<1 | CGMS2>5,NA,CGMS2),
         DM=ifelse(DM>100 | DM<=0,NA,DM),
         DMSPGRAV=ifelse(DMSPGRAV>100 | DMSPGRAV<=0,NA,DMSPGRAV),
         RTWT=ifelse(RTWT==0 | NOHAV==0 | is.na(NOHAV),NA,RTWT),
         SHTWT=ifelse(SHTWT==0 | NOHAV==0 | is.na(NOHAV),NA,SHTWT),
         RTNO=ifelse(RTNO==0 | NOHAV==0 | is.na(NOHAV),NA,RTNO),
         NOHAV=ifelse(NOHAV==0,NA,NOHAV),
         NOHAV=ifelse(NOHAV>42,NA,NOHAV),
         RTNO=ifelse(!RTNO %in% 1:10000,NA,RTNO),
         SPRTPERC=ifelse(SPRTPERC > 1 | SPRTPERC < 0, NA, SPRTPERC), 
         CMD3I=ifelse(CMD3I > 1 | CMD3I < 0, NA, CMD3I), #proportion traits should be between 0 and 1
         CAD6I=ifelse(CAD6I > 1 | CAD6I < 0, NA, CAD6I),
         CBB3I=ifelse(CBB3I > 1 | CBB3I < 0, NA, CBB3I),
         SPRTPERC=ifelse(SPRTPERC > 1 | SPRTPERC < 0, NA, SPRTPERC),
         VIGOR=ifelse(VIGOR!=3 & VIGOR!=5 & VIGOR!=7, NA, VIGOR)) %>%
  
  #fix trials with block and plot number switched: 
  mutate(blockNumber = case_when(studyName == "13uyt20mixedIB" |
                                   studyName == "13uyt15yrt8AIB" |
                                  studyName == "13pyt99highTCIB" |
                                   studyName == "13uyt30whtrtIB" ~ as.character(plotNumber),
                                 TRUE ~ as.character(blockNumber))) %>%
  
  #fix trial with plot number as replicate:
  mutate( replicate = case_when(studyName == "13genpopIB" ~ as.character(blockNumber),
                                TRUE ~ as.character(replicate))) %>%

  #set proper class to variables
  mutate_at(traitnames, as.numeric) %>%
  mutate_at("studyName", as.factor) %>%
  mutate_at("studyYear", as.factor) %>%
  mutate_at("replicate", as.factor) %>%
  mutate_at("blockNumber", as.factor) %>%
  mutate_at("germplasmName", as.factor) %>%
  mutate_at("locationName", as.factor) %>%
  as.data.frame()

#warning about NA generation when setting variables to numeric is okay


saveRDS(trialdatall, file=paste0(project_path, "data/trialdatall_", dataset, ".RDS"))
```

Examine trait distributions and sample sizes of genotyped accessions, subset for the most relevant trials.
```{r, warning=FALSE, message=FALSE}
#filter for the list of genotyped accessions
genotyped_accessions <- samplenames %>%
  gsub("\\.", "-", .) %>% gsub("_A.*", "", .) #remove genotyping sample tag

trialdat_genod <- trialdatall %>% filter(germplasmName %in% genotyped_accessions)


#Trials with iCheck data:
TC_trials <- unique(trialdat_genod %>%  filter(!is.na(.data[[TCtrait]])) %>% dplyr::select("studyName","selStage","germplasmName")) %>% 
  group_by(studyName, selStage) %>%
  summarise(ngenos_TC = n()) %>%
  arrange(-ngenos_TC) 

#Trials with DM data:
DM_trials <- unique(trialdat_genod %>%  filter(!is.na(DM)) %>% dplyr::select("studyName","selStage","germplasmName")) %>% 
  group_by(studyName, selStage) %>%
  summarise(ngenos_DM = n()) %>%
  arrange(-ngenos_DM) 

#Filter trials for at least 3 TCICHK and 10 DM observations on genotyped accessions 
TC_DM_trials <- left_join(TC_trials, DM_trials) %>%
  filter(ngenos_TC > 3 | ngenos_DM > 10)

print(paste0(nrow(TC_DM_trials), " trials with at least 3 TCICHK and 10 DM observations on genotyped accession"))

trials_to_keep <- as.character(TC_DM_trials$studyName)
```


Filter based on non-zero per-trial heritability for carotenoids and dry matter if specified:
Note: this only excludes the trials for which the model didn't fit (singularity issues).
```{r}
if( filter_trial_byH2 == TRUE){
  #Set up storage vector
  TrialH2storage <- data.frame(trials_to_keep, "TC"=NA, "DM"=NA)

  for(i in seq_along(trials_to_keep)){

    #Subset for one trial at a time
    thistrial <- trials_to_keep[i]
    print(i); #print(thistrial)
    thistrialonly <- trialdatall %>% filter(studyName == thistrial) %>% droplevels()
    
    # print(sum(!is.na(thistrialonly$TCICHK)))
    # print(sum(!is.na(thistrialonly$DM)))

    #Check blocking structure
    #print(table(thistrialonly$blockNumber))

    for (j in 1:2){

      traits <- c(TCtrait, "DM")
      traiti <- traits[j]

      fixedformula <- formula(paste0( traiti, "~ 1"))


    trymodelfit <- tryCatch(
      modeli <- asreml(data=thistrialonly,
                       fixedformula,
                       random = ~ germplasmName,
                       trace=F),
          error=function(e) e)

      if(inherits(trymodelfit, "error")){ #if the model didn't fit, put in a placeholder

         TrialH2storage[i, j+1] <- "NA"
      }

      if(!inherits(trymodelfit, "error") && trymodelfit$converge == TRUE){ #if the model successfully fit,

        H2 <- vpredict(modeli, H2 ~ V1 / (V1+V2))
        #print(as.numeric(H2[1]))

        TrialH2storage[i, j+1] <- H2[1]
      }
    } #end trait
  } #end trial


  # Filter for trials with good H2

  PassedTrials <- TrialH2storage %>% filter(TC >= 0.1 & DM >= 0.1)

} else { #keep all trials if filter_trial_byH2 is FALSE

  PassedTrials <- data.frame(trials_to_keep)
}


#filter down to only the more relevant trials
trialdat_subset <- trialdatall %>% filter(studyName %in% PassedTrials$trials_to_keep) %>% droplevels()

#print selected trial info
print(paste0(length(PassedTrials$trials_to_keep), " trials kept"))

trialinfo <- trialdat_subset %>% dplyr::select(studyName, locationName, studyYear, studyDesign, selStage) %>% 
  unique() %>% 
  left_join(TC_trials, by=c("studyName","selStage")) %>%
  left_join(DM_trials, by=c("studyName","selStage")) %>%
  rmarkdown::paged_table()

#make sure all genotypes have phenotype data
#missing accessions: IITA-TMS-IBA190198, IITA-TMS-IBA9100953
print(paste0("missing phenotype data for: ", 
             genotyped_accessions[!genotyped_accessions %in% trialdat_subset$germplasmName]))

#count of trial type 
table(trialinfo$selStage)

#count number of observations of genotypes of interest across traits:
trialinfo %>% dplyr::select(studyName, selStage, ngenos_TC, ngenos_DM)

#selection stage info
selstageinfo <- trialdat_subset %>% dplyr::select(germplasmName, studyYear, selStage) %>% 
  unique() %>%
  filter(germplasmName %in% genotyped_accessions) %>%
  arrange(germplasmName) %>%
  filter(selStage != "NA") %>%
  pivot_longer(selStage, values_drop_na = TRUE) %>%
  pivot_wider(id_cols="germplasmName", names_from=value, values_from=value, values_fn = length, values_fill = 0) %>%
  arrange(UYT, AYT, PYT, CET) %>%
  mutate(stage = case_when(UYT== 0 & AYT == 0 ~ "early", AYT > 2 | UYT > 0 ~ "advanced", T ~ "mid"))
 
table(selstageinfo$stage)/sum(table(selstageinfo$stage)) * 100



saveRDS(trialdat_subset, file=paste0(project_path, "data/trialdata_subset_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))
```



Print table with trait names, abbreviations, and sample sizes
```{r, warning=FALSE, message=FALSE}
trialdat <- trialdat_subset

## Count # of genotypes of interest with data for each trait
traitcounts <- traitnames
for(i in seq_along(traitnames)){
    
  traiti <- traitnames[i]
  
  ngenos_trait <- trialdat %>% filter( !is.na(get(traiti)) ) %>% 
    filter(germplasmName %in% genotyped_accessions) %>%
    dplyr::select(germplasmName) %>% unique() %>% nrow()
  
  traitcounts[i] <- ngenos_trait
  
  #cat(traiti, ngenos_trait, "\n")
}

TraitCounts <- data.frame(TraitAbbrev = traitnames, NGenotypes = as.numeric(traitcounts)) %>%
  filter(NGenotypes > 300)


TraitTable1 <- traitabbrevs
TraitTable1$OntologyID <- str_extract(TraitTable1$TraitName, "CO_334.*")
TraitTable1$TraitName <- gsub("CO_334.*" ,"", TraitTable1$TraitName) %>% gsub("\\."," ", .)

TraitTable1 <- left_join(TraitCounts, TraitTable1, by="TraitAbbrev")

#count accessions with both TC and DM data
TCDMboth <- trialdat %>% filter( !is.na(TCICHK) & !is.na(DM) ) %>% 
    filter(germplasmName %in% genotyped_accessions) %>%
    dplyr::select(germplasmName) %>% unique() %>% nrow()
#363 (but one gets filtered out later in outlier removal, final n = 362)

TraitTable1 <- rbind(TraitTable1, c("TCICHK_DM", TCDMboth, "both TCICHK and DM", "NA"))

  
TraitTable1 %>% rmarkdown::paged_table()
selectedtraits <- TraitTable1$TraitAbbrev[-nrow(TraitTable1)] #each of the major traits except the last TCICHK_DM combo
saveRDS(selectedtraits, file=paste0("./data/selected_traitnames_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))
```


```{r, warning=FALSE, message=FALSE, eval=F}
#make a histogram for each of the major traits
#n = sample size, # of observations

for(i in seq_along(selectedtraits)){ 
    
  traiti <- TraitTable1$TraitAbbrev[i]
  print(traiti)
    
  hist(trialdat[,traiti], main= paste0("n=", nsizes[traiti]), xlab=paste0(traiti))  
}
```

Box-cox correction for normality: 
(note: should I do outlier removal before Box-cox? 1/30/24: decided no, because normality is needed for the linear model that calculates residuals.)
```{r, warning=FALSE, message=FALSE}
trialdat_tfm <- trialdat
  
lambda_values <- data.frame(trait = selectedtraits, lambda=NA)
for(i in seq_along(selectedtraits)){
    
  #for each trait
    traiti <- selectedtraits[i]
    print(traiti)
    traitidata <- trialdat[,traiti]
    traitidata[traitidata <= 0] <- NA #because boxcox requires strictly positive values 
    
  #box-cox transformation to find optimum lambda
    bxcx <- boxcox( lm(traitidata ~ 1), plotit=F)
    lambda <- bxcx$x[which.max(bxcx$y)]

    # apply transformation and write over column
    traiti_transformed <- ((traitidata ^ lambda) - 1) / lambda
    trialdat_tfm[,traiti] <- scale(traiti_transformed) #scale data so traits are on the same scale
    
    #store lambda for downstream un-transformations
    lambda_values[i,"lambda"] <- lambda
}
saveRDS(lambda_values, file=paste0(project_path, "output/Boxcox_lambda_values.RDS"))

### Save checkpoint
saveRDS(trialdat_tfm, file=paste0(project_path, "data/trialdata_tfm_scaled_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))


# check transformations
plot(trialdat$TCICHK, trialdat_tfm$TCICHK, xlab="original", ylab="boxcox transformed", main="TCICHK")
plot(trialdat$DM, trialdat_tfm$DM, xlab="original", ylab="boxcox transformed", main="DM")
plot(trialdat$TCHART, trialdat_tfm$TCHART, xlab="original", ylab="boxcox transformed", main="TCHART")


```


Initial BLUP calculation for identifying and removing outliers based on studentized residuals using asreml
```{r, warning=FALSE, message=FALSE}
#set up blank dataframe to store results
outlier.mat <- data.frame(1:nrow(trialdat_tfm)) 

for(i in seq_along(selectedtraits)){ #go through each trait
  traiti <- selectedtraits[i]; print(traiti)
  
  fixedformula <- formula(paste0( traiti, "~ 1 + locationName:studyYear"))
  
  trymodelfit <- tryCatch( #attempt to fit the model but don't stop the loop if there is an error
    #interaction terms (outer:inner) are nested second within the first
    modi <- asreml(data=trialdat_tfm, fixedformula,
                 random = ~ germplasmName + 
                   germplasmName:locationName:studyYear + studyName:blockNumber, 
                 trace=F, aom=T), 
    error=function(e) e)
  
  if(inherits(trymodelfit, "error") | trymodelfit$converge == FALSE){ #if the model didn't fit, put in a placeholder
    blank <- rep(0, length(nrow(trialdat_tfm)))
    outlier.mat <- cbind(outlier.mat, blank)
  }
  
  if(!inherits(trymodelfit, "error") && trymodelfit$converge == TRUE){ #if the model successfully fit,
    print(paste0(traiti, " model fit"))

    # extract standardized residuals
    stdRes <- resid(modi, type="stdCond")
    resDF <- modi$nedf #residual degrees of freedom
    studentizedRes <- stdRes / sqrt((resDF - stdRes ^ 2) / (resDF - 1))
    
    #set a threshold for outliers
    N <- nrow(modi$mf)
    threshold <- qt(p =  1 - 0.05 / (2 * N), df = (N - 2))
    
    # identify outliers 
    significance <- rep("NA", length(stdRes))
    significance[which(abs(stdRes) <= threshold)] <- 'pass'
    significance[which(abs(stdRes) > threshold)] <- 'outlier_removed'
    print(table(significance))
          
    outlier.mat <- cbind(outlier.mat, significance) #save vector of which points to exclude
  }
    colnames(outlier.mat)[1+i] <- traiti #save the trait name 
}
outlier.mat <- outlier.mat[,-1] #remove first placeholder column


#remove outliers
OutlierRMData <- as.matrix(trialdat_tfm %>% dplyr::select(all_of(selectedtraits)))
#confirm that dimensions match
dim(OutlierRMData) == dim(outlier.mat)
#replace outlier values with NA
OutlierRMData[outlier.mat == "outlier_removed"] <- NA 
trialdat_cleaned <- cbind(trialdat_tfm[,c(1:20)], OutlierRMData)

### Save checkpoint
saveRDS(trialdat_cleaned, file=paste0(project_path, "data/trialdata_scaled_cleaned_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))
```



Calculate BLUPs and H2 with asreml
```{r, warning=FALSE, message=FALSE}
# Re-read dataset if starting here
#dataset <- "CASS" 
trialdat_cleaned <- readRDS(file=paste0(project_path, "data/trialdata_scaled_cleaned_", dataset, "_TCICHK_", "H2filter", "FALSE", ".RDS"))
selectedtraits <- readRDS(file=paste0("./data/selected_traitnames_", dataset, "_TCICHK_", "H2filter", "FALSE", ".RDS"))


#initiate storage 
model_storage <- list() 
BLUP_storage <- data.frame(germplasmName = sort(unique(trialdat_cleaned$germplasmName))) 
H2_storage <- data.frame()
H2_cullis_storage <- data.frame() 


for(i in seq_along(selectedtraits)){ #go through each trait
  
  traiti <- selectedtraits[i]; print(traiti)

  fixedformula <- formula(paste0( traiti, "~ 1 + locationName:studyYear"))
  
  trymodelfit <- tryCatch(
    #note: nested factors are outer:inner
    modi <- asreml(data=trialdat_cleaned, trace=F,
                   fixedformula,
                   random = ~ germplasmName + 
                     germplasmName:locationName:studyYear + studyName:blockNumber,
                   na.action = na.method(y="omit")),
    
    error=function(e) e)
  
  if(!inherits(trymodelfit, "error") && trymodelfit$converge == TRUE){ #if the model converged:
    print(paste0(traiti, " model fit"))
    model_storage[[i]] <- modi #store full model
  

    #extract and de-regress BLUPs
    randomeffs <- summary(modi, coef=TRUE)$coef.random
    Vg <-  summary(modi)$varcomp["germplasmName", "component"]
    
    BLUPs_i <- randomeffs[grepl('germplasmName', rownames(randomeffs)) & !grepl('locationName', rownames(randomeffs)),] %>%
      as.data.frame() %>% 
      mutate(PEV=std.error^2, 
             REL=1-(PEV/Vg),
             drgBLUP=solution/REL) %>% #de-regress BLUPs
      mutate(germplasmName = gsub("germplasmName_","",rownames(.)))
    
    BLUPs_tostore <- BLUPs_i %>% dplyr::select(germplasmName, drgBLUP) %>%
      rename_with(~ paste0(traiti), drgBLUP)
    
    #store de-regressed BLUPs
    BLUP_storage <- left_join(BLUP_storage, BLUPs_tostore, by="germplasmName") 
    
    #calculate heritability
    nLocYearsperG <- trialdat_cleaned %>%
      group_by(germplasmName) %>%
      summarise(n_locyears = n_distinct(locationName:studyYear))
    
    harmonicMeanEnv <- 1 / mean(1 / nLocYearsperG$n_locyears)

    nPlotsPerG <- trialdat_cleaned %>%
      group_by(germplasmName) %>%
      summarise(n_rep = n_distinct(locationName:studyYear:replicate))
    
    harmonicMeanPlots <- 1 / mean(1 / nPlotsPerG$n_rep)
    
    Vg <-  summary(modi)$varcomp["germplasmName", "component"]
    Vgly <- summary(modi)$varcomp["germplasmName:locationName:studyYear", "component"]
    Ve <- summary(modi)$varcomp["units!R", "component"]

    H2 <- Vg / sum(Vg + (Vgly/harmonicMeanEnv) + (Ve/harmonicMeanPlots))
    H2_storage <- rbind(H2_storage, H2) #store Heritability estimate
     
    
    #calculate heritability with Cullis 2006 method
    H2_cullis <- 1 - ((mean(BLUPs_i$PEV))/Vg)
    H2_cullis_storage <- rbind(H2_cullis_storage, H2_cullis)
    
  } 
  else{ #if there was a problem with the model, use NA as placeholders in storage vectors
    print(paste0(traiti, " model failed to converge")) 
    model_storage[[i]] <- NA
     BLUP_storage <- cbind(BLUP_storage, rep(NA, length(unique(trialdat_cleaned$germplasmName))))
     colnames(BLUP_storage)[1+i] <- traiti
     H2_storage <- rbind(H2_storage, c(NA,NA))
     H2_cullis_storage <- rbind(H2_cullis_storage, "NA")
    }
}
names(model_storage) <- selectedtraits

rownames(H2_storage) <- selectedtraits #reattach the trait names
colnames(H2_storage) <- "H2, var. comp. method"

rownames(H2_cullis_storage) <- selectedtraits #reattach the trait names
colnames(H2_cullis_storage) <- "H2, Cullis 2006 method"


### Save BLUPs
# saveRDS(BLUP_storage, file=paste0(project_path, "data/alltrait_scaled_drgBLUPs_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))
# saveRDS(H2_storage, file=paste0(project_path, "data/alltrait_scaled_H2_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))
# saveRDS(H2_cullis_storage, file=paste0(project_path, "data/alltrait_scaled_H2_cullis_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2,".RDS"))
# saveRDS(model_storage, file=paste0(project_path, "data/BLUPmodel_scaled_storage_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))

```



Examine phenotypic data:
```{r, warning=FALSE, message=FALSE}
H2_storage<- readRDS(file=paste0(project_path, "data/alltrait_scaled_H2_CASS_TCICHK_H2filter", filter_trial_byH2, ".RDS"))
BLUP_storage <- readRDS(file=paste0(project_path, "data/alltrait_scaled_drgBLUPs_CASS_TCICHK_H2filter", filter_trial_byH2, ".RDS"))
model_storage <- readRDS(file=paste0(project_path, "data/BLUPmodel_scaled_storage_CASS_TCICHK_H2filter", filter_trial_byH2, ".RDS"))
trialdat_cleaned <- readRDS(file=paste0(project_path, "data/trialdata_scaled_cleaned_CASS_TCICHK_H2filter", filter_trial_byH2, ".RDS"))
genotyped_accessions <- read.table("./data/genotypenames_CASS.txt")[,1] %>%
  gsub("\\.", "-", .) %>% gsub("_A.*", "", .)
traitsofinterest <- c("TCICHK", "TCHART", "DM", "DYLD", "RTWT", "HI", "SPRTPERC", "FRYLD", "CMD3S", "CMD3I")


#print heritability -- for table 1
H2_storage %>% round(2) %>% rmarkdown::paged_table()

#print number of observations
for(i in 1:length(selectedtraits)){
 cat(selectedtraits[i], 
            nrow(trialdat_cleaned %>% filter(!is.na(.data[[selectedtraits[i]]])) ), "\n" )
}


#Calculate pairwise correlations between trait BLUPs
BLUP_storage_genod <- BLUP_storage %>% filter(germplasmName %in% genotyped_accessions)
cormat <- round(cor(BLUP_storage_genod[,-1], use="pairwise.complete.obs"), 3)
cormat[upper.tri(cormat)] <- NA #remove upper triangle

#get p-values for correlations
cor.test.p <- function(x){
    FUN <- function(x, y) cor.test(x, y)[["p.value"]]
    z <- outer(
      colnames(x), 
      colnames(x), 
      Vectorize(function(i,j) FUN(x[,i], x[,j]))
    )
    dimnames(z) <- list(colnames(x), colnames(x))
    z
}
cormat_pvals <- cor.test.p(BLUP_storage_genod[,-1])
cormat_pvals[upper.tri(cormat_pvals)] <- NA #remove upper triangle

melted_pvals <- reshape2::melt(cormat_pvals) %>% filter(!is.na(value)) %>%
  mutate(pvalue = signif(value, 3)) %>% dplyr::select(-value) %>%
  filter(Var1 %in% traitsofinterest & Var2 %in% traitsofinterest)

traitsofinterest <- c("TCICHK", "TCHART", "DM", "DYLD", "RTWT", "HI", "SPRTPERC", "FRYLD", "CMD3S", "CMD3I")

melted_cormat <- reshape2::melt(cormat) %>% filter(!is.na(value)) %>%
  mutate(Rlabel = round(value, 2)) %>% 
  filter(Var1 %in% traitsofinterest & Var2 %in% traitsofinterest) %>%
  left_join(melted_pvals) %>%
  mutate(sig = as.factor(case_when(Var1 == Var2 ~ "plain",
                                   pvalue < 0.05 ~ "bold", 
                                   T ~ "plain")))

#heatmap of trait correlation matrix
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  #geom_tile(aes(color = sig, width=0.95, height=0.95), size=0.8) +
  #scale_color_manual(values=c("white","black"), guide="none") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\ncorrelation\ncoefficient (r)") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
     hjust = 1))+
  coord_fixed() +
  xlab("") +
  ylab("") +
  geom_text(aes(label=Rlabel, fontface=sig), size=3, parse=F) +
  ggtitle(paste0("Genetic correlations between traits")) +
  theme(panel.grid = element_blank())

ggsave(file=paste0(project_path, "figures/TraitScaledCorrMatrix_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".jpg"), bg="white")


#Phenotypic correlation between TC and DM
phenocor <- cor.test(trialdat_cleaned$TCICHK, trialdat_cleaned$DM, na.rm=T)
phenocor
phenocor$p.value
#plot(trialdat_cleaned$TCICHK, trialdat_cleaned$DM)
#cor.test(trialdat_cleaned$TCHART, trialdat_cleaned$DM, na.rm=T)

#Genetic correlation between TC and DM in full dataset
genocor <- cor.test(BLUP_storage$TCICHK, BLUP_storage$DM, na.rm=T)
genocor
genocor$p.value
# there is a stronger negative genetic correlation than there is phenotypic

#Genetic correlation between TC and DM in genotyped subset
TCICHK_DM_cor <- cor.test(BLUP_storage_genod$TCICHK, BLUP_storage_genod$DM, na.rm=T)
TCICHK_DM_cor
TCICHK_DM_cor$p.value


ggplot(data=BLUP_storage_genod, aes(TCICHK, DM)) +
  geom_point() +
  ggtitle("Genetic correlation between TCICHK and DM") +
  geom_smooth(method='lm', formula= y~x, se=F, lty=2, size=0.5) +
  annotate("text", x=1.25 , y= 1.75, label= paste0("r = ", round(TCICHK_DM_cor$estimate,2))) +
  annotate("text", x=1.25 , y= 1.5, label=paste0("p = ", signif(TCICHK_DM_cor$p.value, digits=3))) +
  xlab("TCICHK BLUP") +
  ylab("DM BLUP")

#ggsave("~/Desktop/images_dwnld/TCDMcorplot.png", bg="white", width=6, height=5)

ggsave(file=paste0(project_path, "figures/TCDM_scaled_cor_plot_all", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".jpg"))



#Genetic correlation between TC and DM in yellow-only dataset
ggplot(data=BLUP_storage %>% filter(TCICHK>-1.25), aes(TCICHK, DM)) +
  geom_point() +
  ggtitle("Genetic correlation between TC and DM BLUPs only in 'yellow' clones") +
  geom_smooth(method='lm', formula= y~x, se=F, lty=2, size=0.5)
ggsave(file=paste0(project_path, "figures/TCDM_scaled_cor_plot_ylw", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".jpg"))



#Calculate pairwise correlations between residuals
TCICHK_resids <- model_storage[[1]]$residuals
DM_resids <- model_storage[[3]]$residuals

cor.test(TCICHK_resids, DM_resids)
#residuals are also (very) slightly negatively correlated

ggplot(data=as.data.frame(TCICHK_resids, DM_resids), 
       aes(TCICHK_resids, DM_resids)) +
  geom_point() +
  ggtitle("Correlation between TC and DM residuals") +
  geom_smooth(method='lm', formula= y~x, se=F, lty=2, size=0.5)

#Check that the residuals are normally distributed -- looks good
hist(TCICHK_resids, main="TC residuals", xlab="residuals")
hist(DM_resids, main="DM residuals", xlab="residuals")
```





How many white vs yellow accessions are there?
```{r}
dataset <- "CASS"
filter_trial_byH2 <- TRUE
TCtrait <- "TCICHK"


trialdat_cleaned <- readRDS(file=paste0(project_path, "data/trialdata_cleaned_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))

trialdat_subset <- readRDS(file=paste0(project_path, "data/trialdata_subset_", dataset, "_", TCtrait, "_", "H2filter", filter_trial_byH2, ".RDS"))

BLUPs <- read.table(file=paste0(project_path, "data/BLUPmatrix_named_", dataset, ".txt"), header=T) %>%
  mutate(germplasmName = as.factor(sample_names))

trialdat_subset_colorsum <- trialdat_subset %>% filter(germplasmName %in% BLUPs$sample_names) %>%
  group_by(germplasmName) %>%
  summarise(meanTCHART = mean(TCHART, na.rm=T), meanTCICHK = mean(TCICHK, na.rm=T)) %>%
  mutate(category = case_when(meanTCICHK >=4 ~ "yellow", 
                              meanTCHART > 1.5 & is.na(meanTCICHK) ~ "yellow",
                              is.na(meanTCHART) & meanTCICHK >=4 ~ "yellow", 
                              meanTCHART > 1.5 | meanTCICHK < 4 ~ "white", T~"NA"))

# save names of yellow accessions
yellow_only <- trialdat_subset_colorsum %>% filter(category == "yellow")
saveRDS(yellow_only, file=paste0(project_path, "yellowSubpop/yellowNamesSummary.RDS"))

ggplot(trialdat_subset_colorsum, aes(meanTCHART, meanTCICHK)) +
  geom_point(aes(fill=meanTCICHK), pch=21, color="dark gray", size=2) +
  scale_fill_gradient2(low="white", mid="#ffec5c", high="dark orange", midpoint=15)

#saveRDS(trialdat_subset_colorsum, file=paste0(project_path, "data/genotyped_accessions_color_categorization_CASS.RDS"))
```



