---
title: "submitted_version_mvMLMM"
output: html_document
date: "2025-07-15"
---


```{r}
setwd("/workdir/ssv42/mvGWAS/")
library(tidyverse)
project_path <- "/workdir/ssv42/mvGWAS/"
```

Set-up
(Note: copy and paste into terminal R session so Rstudio remains free)
```{r}
library(tidyverse); library(asreml); library(aod)
project_path <- "/workdir/ssv42/mvGWAS/"
dataset <- "CASS"
ncores <- 60
  
### DEBUG settings -- test run one SNP model
DEBUG <- TRUE

#read in original drgBLUP phenotype data
BLUPs_data <- readRDS(file=paste0(project_path, "data/BLUPs_scaled_and_hap_data_", dataset, ".RDS")) 
  
# Read in revised genotype matrix 
genomat_SNPs <- readRDS(file="./data/genomat_rev.RDS")
#dim(genomat_SNPs) #377 x 43720
  
# Filter for common accessions
common_acc <- intersect(rownames(genomat_SNPs), BLUPs_data$germplasmName)
genomat_SNPs <- genomat_SNPs[common_acc,]
BLUPs_data <- BLUPs_data %>% filter(germplasmName %in% common_acc)
      
# Get fresh relationship matrix
# #just for the purposes of calculating G matrix: round imputed values to integers
# genomat_rounded <- round(genomat_SNPs)
# GMatrix <- ASRgenomics::G.matrix(M=as.matrix(genomat_rounded), method="VanRaden")[[1]]

# # Blend GRM to improve diagonal values
# Gblend <- ASRgenomics::G.tuneup(GMatrix, blend=TRUE, pblend=0.05)$Gb
# checkG <- ASRgenomics::kinship.diagnostics(Gblend)

# # Invert GRM
# Ginverse <- ASRgenomics::G.inverse(Gblend, sparseform=FALSE)$Ginv
# saveRDS(Ginverse, file=paste0(project_path, "data/Ginverse_revisedimputation.RDS"))
      
Ginverse <- readRDS(file=paste0(project_path, "data/Ginverse_revisedimputation.RDS"))
Gsampleorder <- rownames(Ginverse) #make sure names are the same order
      
      
# Add computational group sections to SNPs
genomat_t <- t(genomat_SNPs) %>% as.data.frame()
genomat_t$SNP <- rownames(genomat_t)
nSNPs <- dim(genomat_t)[1]
genomat_t$compute_group <- as.numeric(cut_number(c(1:nSNPs), ncores))
#saveRDS(genomat_t, file=paste0(project_path, "data/genomat_t_revised.RDS"))
      
      
# Read in imputed PSY2 KASP marker covariate
PSY2_imputed_rev <- readRDS(file=paste0(project_path, "data/PSY2KASP_revimputed_unrounded.RDS"))
      # Get original PSY2 KASP marker covariate
      # genomat_KASP_OG <- readRDS(file=paste0(project_path, "data/genomat_plusKASP_filtered.RDS"))
      #  PSY2marker_OG <- as.data.frame(t(genomat_KASP_OG[genomat_KASP_OG$SNP == "S1_24155522",] %>%
      #   dplyr::select(-c("SNP","ALT","REF","compute_group"))))
      #  colnames(PSY2marker_OG) <- "S1_24155522"
      #  PSY2marker_OG$germplasmName <- rownames(PSY2marker_OG)
      #  PSY2marker_OG$S1_24155522_dom <- ifelse(PSY2marker_OG$S1_24155522 == 1, 1, 0) 
    
#Get PC covariates
PCcovariates <- readRDS("/workdir/ssv42/mvGWAS/data/PCcovariates_rev.RDS") %>%
  as.data.frame() %>% rownames_to_column("germplasmName")
      # Get original PC covariates
       # PCaccessionnames <- read.table(file= paste0(project_path, "data/", dataset, "_accessionnames.txt"))[,1]
       # dosagesPC <- as.data.frame(readRDS(file=paste0(project_path,
       #"data/DosageMatrix_PCA_CASS_PCmatrix_copy.RDS")))[,1:5]
       # dosagesPC$germplasmName <- PCaccessionnames
      
#Get weights
BLUPs_revised <- readRDS(file=paste0(project_path,
                                     "data/alltrait_scaled_drgBLUPs_rev_CASS_TCICHK_H2filterFALSE.RDS")) %>%
  filter(germplasmName %in% Gsampleorder) %>%
  mutate(acc_weight = rowMeans(cbind(TCICHK_weight, DM_weight), na.rm = TRUE)) %>% 
  #use mean of the two traits for accession-level weights 
  mutate(acc_weight = case_when(is.na(acc_weight) ~ min(acc_weight, na.rm=T), T ~ acc_weight))
  #fill any NAs with the lowest weight
BLUP_weights <- BLUPs_revised %>% dplyr::select(germplasmName, TCICHK_weight, DM_weight, acc_weight)
      
#Merge BLUP and PSY2 marker data together, ensuring correct order of sample names
BLUPs_PSY2_data <- left_join(data.frame(germplasmName = Gsampleorder), BLUPs_data, by="germplasmName") %>%
  left_join(PSY2_imputed_rev, by="germplasmName") %>%
  left_join(PCcovariates, by="germplasmName") %>%
  left_join(BLUP_weights, by="germplasmName") %>%
  mutate(germplasmName = as.factor(germplasmName))

#saveRDS(BLUPs_PSY2_data, file=paste0(project_path, "data/BLUPs_PSY2_revised_data_0716.RDS"))
      
#confirm correct order
all(BLUPs_PSY2_data$germplasmName == Gsampleorder)
length(BLUPs_PSY2_data$germplasmName) == 376
nAccessions <- length(BLUPs_PSY2_data$germplasmName) 
      
# Get starting values to help model
T1 <- asreml(data = BLUPs_PSY2_data, 
                 fixed = TCICHK ~ 1 + S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5, 
                 random = ~ vm(germplasmName, Ginverse),
                 residual = ~units)
      
Vg_TCICHK <- summary(T1)$varcomp$component[1]
      
T2 <- asreml(data = BLUPs_PSY2_data, 
                   fixed = DM ~ 1 + S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5, 
                   random = ~ vm(germplasmName, Ginverse),
                   residual = ~units)
Vg_DM <- summary(T2)$varcomp$component[1]
      
Cov_TCICHK_DM <- cor(BLUPs_PSY2_data$TCICHK, BLUPs_PSY2_data$DM, use='complete.obs')
initG <- c(Cov_TCICHK_DM, Vg_TCICHK, Vg_DM) #trait *correlation*, TCICHK gen variance, DM gen variance

#saveRDS(initG, file=paste0(project_path, "data/initG_revisedimputation.RDS"))
    

```

Define mvGWAS function
```{r}

SNP_GWAS_PCs_function <- function(section){
  
  #load packages
  library(tidyverse); library(asreml); library(aod)
  
  #select SNPs in this compute group
  section_SNPs <- genomat_t %>% filter(compute_group == section)

  
  # Loop across all "rest of genome" SNPs 
  #storage for SNP effects
  all_SNP_results_df <- as.data.frame(matrix(nrow = nrow(section_SNPs), ncol = 18))
  colnames(all_SNP_results_df) <- c("level", "status", 
                                    "BetaSNP_TCICHK", "BetaSNP_DM", "BetaSNP_TCICHK.se", "BetaSNP_DM.se", "BetaSNP_TCICHKDM.cov", 
                                   "PSY2_TCICHK", "PSY2_DM", "PSY2_TCICHK.se", "PSY2_DM.se",
                                    # "PSY2_dom_TCICHK", "PSY2_dom_TCICHK.se", "PSY2_dom_DM", "PSY2_dom_DM.se", 
                                    "chi2", "df", "P", "resid_gen_cov", "resid_gen_cov.se", "error_cov", "error_cov.se")
  
  #storage for model objects
  #model_storage <- list()

  SNPstotest <- seq_along(section_SNPs$SNP)
  if(DEBUG == TRUE){ SNPstotest <- 1 }
  
  for(i in SNPstotest){
    
    #debug note: chr 1 peak is at i = 292 
    SNPi <- section_SNPs$SNP[i]
    SNPidata <- as.data.frame(t(section_SNPs[i,1:nAccessions]))
    colnames(SNPidata) <- paste0(SNPi)
    SNPidata$germplasmName <- rownames(SNPidata)

    modeldata <- left_join(BLUPs_PSY2_data, SNPidata, by="germplasmName") %>% 
      mutate(germplasmName = as.factor(germplasmName))
    #modeldata <<- modeldata
    
    
    # DEBUG
    # test base model
    # basemodel <- asreml(data = modeldata,
    #     formula(paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
    #      random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
    #     residual = ~units:corgh(trait), maxit=20)
    
    # test weighted model -- doesn't work (Sigma2 goes to 0, LogLik goes to -inf)
    # weightsmodel <- asreml(data = modeldata,
    #     formula(paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
    #      random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
    #     residual = ~units:corgh(trait), maxit=20, weights=acc_weight_inv)
  
    #model calculating SNP effect on each trait
    SNPimodel <- asreml(data = modeldata,
        formula(paste0("cbind(TCICHK, DM) ~ trait + trait:", SNPi, 
                       " + trait:S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
        residual = ~units:us(trait), maxit=20, weights=acc_weight)
    
    model_varcomp_i <- summary(SNPimodel)$varcomp
    PSY2SNPestimates <- summary(SNPimodel, coef=T)$coef.fixed[1:2,1:2]
  
    #save model
    #SNPimodel <<- SNPimodel
    #model_storage[[i]] <<- SNPimodel

    #predict SNP effects
    SNPlevel <- list(1)
    names(SNPlevel) <- SNPi
    predictSNPeffects <- predict.asreml(SNPimodel, vcov=T, 
                                        classify=paste0("trait:", SNPi), 
                                        only=paste0("trait:", SNPi), 
                                        levels=SNPlevel)
    
    Betas_i <- pivot_wider(predictSNPeffects$pvals, names_from = "trait", 
                           values_from=c("predicted.value", "std.error"))
    Cov_i <- predictSNPeffects$vcov
    
    #Bivariate Wald test:
    WaldTest <- aod::wald.test(Sigma=predictSNPeffects$vcov,
                               b=predictSNPeffects$pvals$predicted.value, Terms=1:2)
    
    
    #Save results
    all_SNP_results_df[i,] <- cbind(Betas_i, Cov_i[1,2], 
                                    t(PSY2SNPestimates[,1]), t(PSY2SNPestimates[,2]),
                                    t(WaldTest$result$chi), model_varcomp_i[1,1:2], 
                                    model_varcomp_i["units:trait!trait!DM:!trait!TCICHK.cor",1:2])

    rownames(all_SNP_results_df)[i] <- colnames(Betas_i)[1]
  
  saveRDS(all_SNP_results_df, file=paste0(project_path, 
                          "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_weighted_section_", 
                           section, "_", format(Sys.time(), "%m%d"), ".RDS"))
 
  cat(paste0(SNPi, ", ", i, " of ", length(section_SNPs$SNP), " in section ", section, "\n"),
      file = paste0(project_path, "output/mvGWAS_asreml_output/progressbar/unrounded_OGBLUPs_weighted_run_071625.txt"), append=T)

  } #end SNP

} #end compute section
```

Run function on parallel cores
```{r}
library(parallel)
DEBUG <- FALSE

#Run asreml GWAS function over multiple cores:
sections <- 1:ncores

# export functions and data to cluster
GWASfunctionvariables <- c("SNP_GWAS_PCs_function", "genomat_t", "Ginverse", "initG","BLUPs_PSY2_data", 
                           "project_path", "ncores", "DEBUG", "nAccessions")
  
# create cluster object
cl <- makeCluster(ncores)
clusterExport(cl, varlist = GWASfunctionvariables)

# test each SNP in chromosome 
results <- parSapply(cl, sections, SNP_GWAS_PCs_function)

# close cluster object
stopCluster(cl)
```

#### Original BLUPs, re-imputed unrounded genotype data, unweighted
Combine (unweighted) results from the parallel computation sections: 
```{r}
project_path <- "/workdir/ssv42/mvGWAS/"
dataset <- "CASS"
ncores <- 60
lastrundate <- "0716" # may need a tryCatch if sections finished on different dates

all_SNP_results_df <- data.frame()
for (i in 1:ncores){
  section <- i
  print(i)
  
  lookupresults <- tryCatch(
    SNPresults_i <- readRDS(file=paste0(project_path,
               "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_section_", 
               section, "_", lastrundate, ".RDS")),
    error = function(e) e)

  # if(inherits(lookupresults, "error") ){
  #   print("trying previous day")
  #   lastrundate <- "1706"
  #    SNPresults_i <- readRDS(file=paste0(project_path,
  #              "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_section_", 
  #              section, "_", lastrundate, ".RDS"))
  # }

  all_SNP_results_df <- rbind(all_SNP_results_df, SNPresults_i)
}

# extract SNP information
all_SNP_results_df$SNP_ID <- rownames(all_SNP_results_df)
all_SNP_results_df$CHROM <- gsub("_[0-9].*", "",  rownames(all_SNP_results_df)) %>% gsub("S","",.)
all_SNP_results_df$POS <- gsub("S[0-9]*_", "",  rownames(all_SNP_results_df))


# Calculate significance of residual genetic covariance:
#all_SNP_results_df$resid_gen_cov.z <- all_SNP_results_df$resid_gen_cov / all_SNP_results_df$resid_gen_cov.se
#Calculate p-value from Z-score
#all_SNP_results_df$resid_gen_cov.p <- pnorm(all_SNP_results_df$resid_gen_cov.z, lower.tail = TRUE)


saveRDS(all_SNP_results_df, file=paste0(project_path,
             "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_allSNPresults_",
             format(Sys.time(), "%m%d"), ".RDS"))

```

## Examine the weighted results
Make Q-Q plots
```{r}
all_SNP_weighted_results_df <- readRDS(file=paste0(project_path,
            "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_weighted_allSNPresults_0716.RDS"))

all_SNP_results_MVWald <- all_SNP_weighted_results_df

png(paste0(project_path, "figures/mvGWAS_wPSY2KASPand5PCs_unroundedgenos_weighted_QQplot_CASS_",
                    format(Sys.time(), "%m%d"), ".png"), 
        width=6, height=6, res=250, units="in")
 
qqman::qq(all_SNP_results_MVWald$P, main="Q-Q plot of weighted mvGWAS with reimputed genotypes")

dev.off()

# check the distribution of P-values
ggplot(all_SNP_results_MVWald, aes( BetaSNP_TCICHK/BetaSNP_TCICHK.se, 
                                    BetaSNP_DM/BetaSNP_DM.se, 
                                    color=-log10(P))) +
  ggtitle("w PSY2 covariate and PCs, weighted") +
  geom_point()


# check the distribution of the polygenic covariance between TC and DM -- is it still negative?
hist(all_SNP_results_MVWald$resid_gen_cov, main="residual genetic covariance between TC/DM \nafter accounting for PSY2 and SNP, weighted", xlab="covariance estimate")

range(all_SNP_results_MVWald$resid_gen_cov, na.rm=T)

```

Make Manhattan plot 
```{r}
all_SNP_results_MVWald$CHROM <- gsub("_[0-9]*","",all_SNP_results_MVWald$SNP_ID) %>% gsub("S","",.)
all_SNP_results_Bivar <- all_SNP_results_MVWald

# Set significance cut-off: Gao-modified Bonferroni
Gao.cutoff <- 1.367e-05

all_SNP_results_Bivar$POS <- as.numeric(all_SNP_results_Bivar$POS)
MVresults_sort <- all_SNP_results_Bivar %>% arrange(as.numeric(CHROM), as.numeric(POS))

#order factor levels so they display in order of chromosome
MVresults_sort$SNP_ID <- factor(MVresults_sort$SNP_ID, levels = MVresults_sort$SNP_ID)
MVresults_sort$CHROM <- as.factor(MVresults_sort$CHROM)
MVresults_sort$CHROM <- factor(MVresults_sort$CHROM, levels = as.character(c(1:18)))


#set color scheme
colPalette <- rep(c("blue4","pink3"),9)

## Plot Manhattan with bivariate Wald test p-values
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color=CHROM), size=0.5) +
  scale_color_manual(values=colPalette) +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes, weighted")) +
  guides(color="none")+
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(legend.position = "none",
          strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP
#PSY2 v6 is 24153419 - 24156720

ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_unrounded_weighted_", format(Sys.time(), "%m%d_%H%M"), ".jpg"), width=12, height=5)


#for poster:
ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_unrounded_weighted_CASS_", format(Sys.time(), "%m%d_%H%M"), "_wide.jpg"), width=12, height=3)


saveRDS(MVresults_sort, file=paste0(project_path, "output/MVresults_sorted_PSY2KASP_5PCs_revision_unroundedgenotypes_weighted.RDS"))

```

Color Manhattan plot by SNP covariance
```{r}
hist(MVresults_sort$BetaSNP_TCICHKDM.cov, main="Distribution of SNP TCICHK/DM effects covariance", xlab="SNP effect covariance")
hist((MVresults_sort %>% filter(P < 0.001))$BetaSNP_TCICHKDM.cov, main="Distribution of top SNP TCICHK/DM effects covariance for top SNPs", xlab="SNP effect covariance")


MVresults_sort$effectdir <- case_when(
      MVresults_sort$P > 0.001 ~ "NA",
      MVresults_sort$BetaSNP_TCICHK * MVresults_sort$BetaSNP_DM >= 0 ~ "same",
      MVresults_sort$BetaSNP_TCICHK * MVresults_sort$BetaSNP_DM < 0 ~ "different")

colPalette2 <- c("#00BFC4","gray", "#F8766D")

## Plot Manhattan with covariance colored
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color=effectdir), size=0.5) +
  scale_color_manual(values=colPalette2) +
  #scale_color_gradient2(low = "blue", mid="gray", high = "red", midpoint = 0, name="SNP effect covariance") +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes, weighted")) +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  geom_vline(aes(xintercept=0), lty="dashed", color="gray", linewidth=0.5) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP
#PSY2 v6 is 24153419 - 24156720

ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_revised_unrounded_weighted_weffectdir_", format(Sys.time(), "%m%d_%H%M"), ".jpg"), width=12, height=5)

ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_revised_unrounded_weighted_weffectdir_", format(Sys.time(), "%m%d_%H%M"), "_wide.jpg"), width=8, height=2)


table((MVresults_sort %>% filter(P < 0.001))$effectdir)

table((MVresults_sort %>% filter(P < Gao.cutoff))$effectdir)


## Plot Manhattan with SNP effect error covariance colored
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color=BetaSNP_TCICHKDM.cov), size=0.5) +
  #scale_color_manual(values=colPalette2) +
  scale_color_gradient2(low = "blue", mid="gray", high = "red", midpoint = 0, name="SNP effect covariance") +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes, weighted")) +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  geom_vline(aes(xintercept=0), lty="dashed", color="gray", linewidth=0.5) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP


## Plot Manhattan with residual (polygenic) genetic covariance colored
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color= resid_gen_cov), size=0.5) +
  scale_color_manual(values=colPalette2) +
  scale_color_gradient2(high = "blue", low="gray", mid = "lightblue", midpoint = -0.4, limits=c(-0.6,-0.19),
                        name="residual \ngenetic \ncovariance") +
 #scale_color_gradient(high = "blue", low="gray", name="residual polygenic covariance") +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes, weighted")) +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  #geom_vline(aes(xintercept=0), lty="dashed", color="gray", linewidth=0.5) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP
#PSY2 v6 is 24153419 - 24156720


#note: the residual genetic covariance gets smaller (less negative) when the SNP effects the traits in different directions (i.e. the SNP is explaining some of the covariance)
#hist((MVresults_sort %>% filter(effectdir=="same"))$TCDMgencov)
#hist((MVresults_sort %>% filter(effectdir=="different"))$TCDMgencov)
```

SNP contribution to negative genetic covariance:
```{r}
BLUPs_PSY2_data <- readRDS(file=paste0(project_path, "data/BLUPs_PSY2_revised_data_0716.RDS"))
initG <- readRDS(file=paste0(project_path, "data/initG_revisedimputation.RDS"))
Ginverse <- readRDS(file=paste0(project_path, "data/Ginverse_revisedimputation.RDS"))

#Calculate total genetic variance/covariance with no other SNPs or PSY2
#model calculating SNP effect on each trait
nullmodel <- asreml(data = BLUPs_PSY2_data,
        formula(paste0("cbind(TCICHK, DM) ~ trait  + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
        residual = ~units:us(trait), weights=acc_weight, maxit=20)
nullmodel_varcomp <- summary(nullmodel)$varcomp
    
nullmodel_TCICHKv <- nullmodel_varcomp["trait:vm(germplasmName, Ginverse)!trait_TCICHK", 1] +
  nullmodel_varcomp["units:trait!trait_TCICHK:TCICHK", 1]
nullmodel_DMv <- nullmodel_varcomp["trait:vm(germplasmName, Ginverse)!trait_DM", 1] +
  nullmodel_varcomp["units:trait!trait_DM:DM", 1]
nullmodel_cov <- nullmodel_varcomp["trait:vm(germplasmName, Ginverse)!trait!DM:!trait!TCICHK.cor", 1] + 
  nullmodel_varcomp["units:trait!trait_DM:TCICHK", 1]

# null model with PSY2 
nullmodel_PSY2 <- asreml(data = BLUPs_PSY2_data,
        formula(paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
        residual = ~units:us(trait), weights=acc_weight, maxit=20)
nullmodel_PSY2_varcomp <- summary(nullmodel_PSY2)$varcomp

nullmodel_PSY2_TCICHKv <- nullmodel_PSY2_varcomp["trait:vm(germplasmName, Ginverse)!trait_TCICHK", 1] + 
  nullmodel_PSY2_varcomp["units:trait!trait_TCICHK:TCICHK", 1]
nullmodel_PSY2_DMv <- nullmodel_PSY2_varcomp["trait:vm(germplasmName, Ginverse)!trait_DM", 1] + 
  nullmodel_PSY2_varcomp["units:trait!trait_DM:DM", 1]
nullmodel_PSY2_cov <- nullmodel_PSY2_varcomp["trait:vm(germplasmName, Ginverse)!trait!DM:!trait!TCICHK.cor", 1] + 
  nullmodel_PSY2_varcomp["units:trait!trait_DM:TCICHK", 1]



#difference in polygenic covariance
# if the negative covariance goes up (more positive), then the SNP is contributing to the negative covariance
MVresults_sort$contributiontocov <- nullmodel_gcov - MVresults_sort$resid_gen_cov

## Plot Manhattan with residual (polygenic) genetic covariance colored
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color= contributiontocov), size=0.5) +
  #scale_color_manual(values=colPalette2) +
  scale_color_gradient2(high = "red", low="blue", mid = "gray", midpoint = 0.02,
                        name="contribution to \ngenetic covariance \nbetween TC and DM") +
 #scale_color_gradient(high = "blue", low="gray", name="residual polygenic covariance") +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes, weighted")) +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  #geom_vline(aes(xintercept=0), lty="dashed", color="gray", linewidth=0.5) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP
#PSY2 v6 is 24153419 - 24156720



ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_contributiontogencov_revised_unrounded_weighted_", format(Sys.time(), "%m%d_%H%M"), "_wide.jpg"), width=12, height=3)
```

Examine LD structure around top SNPs
```{r}
#note: starts with MVresults_sort from chunk above
all_SNP_results_df <- MVresults_sort
genomat_t <- readRDS(file=paste0(project_path, "data/genomat_t_revised.RDS"))

#set relaxed threshold 
relaxed.cutoff <- 0.001
#Meffs <- readRDS(file=paste0(project_path, "output/Meff_rev_estimates_perchr.RDS"))
#sum(Meffs) #=594
#old way: #relaxed.cutoff <- round(0.25/sum(Meffs),7)  #0.0004
#print(relaxed.cutoff)

all_SNP_results_df$POS <- as.numeric(all_SNP_results_df$POS)
topSNPs <- all_SNP_results_df %>% filter(P < relaxed.cutoff) #relaxed threshold

genomat_ROG <- genomat_t %>%
   mutate(CHROM= as.integer(gsub("S", "", SNP) %>% gsub("_[0-9].*", "", .)),
          POS= as.numeric(gsub("^S.*_", "", SNP)))

#flag peaks and calculate LD -- takes a minute to run
genomat_peaks_allchrom <- data.frame()
for(i in 1:18) {
  print(i)
  genomat_peaks_i <- genomat_ROG %>% filter(CHROM == i) 

  topSNPs_chri <- topSNPs %>% filter(CHROM == i) %>% arrange(P)
  peak1 <- topSNPs_chri[1,]

  #calculate LD between each SNP and the peak 1 SNP
  if(nrow(peak1) != 0 & !is.na(peak1$SNP_ID)){
    LDmatrix_peak1 <- genomat_peaks_i %>% dplyr::select(-c(SNP, compute_group, CHROM, POS)) %>% #drop metadata non-numeric columns
      t() %>% cor() #calculate correlation matrix
    LDwithpeak1 <- LDmatrix_peak1[,paste0(peak1$SNP_ID)] #pull out correlation with peak1 SNP
    genomat_peaks_i$LDpeak1 <- LDwithpeak1
    
    genomat_peaks_i <- genomat_peaks_i %>% mutate(distance_peak1 = POS - peak1$POS, peak1_SNP = peak1$SNP_ID)
  } else{genomat_peaks_i <- genomat_peaks_i %>% mutate(LDpeak1 = NA, distance_peak1 = NA, peak1_SNP = NA)}

  #check if there is another peak on this chromosome and repeat if so
  peak1_region <- c( peak1$POS - 2000000 , peak1$POS + 2000000)  
  genomat_around_peak1 <- genomat_peaks_i %>% 
       filter(POS > peak1_region[1] & POS < peak1_region[2])
  
  otherpeaks <- topSNPs_chri[!topSNPs_chri$SNP_ID %in% genomat_around_peak1$SNP,]
  peak2 <- otherpeaks %>% arrange(P) %>% head(n=1)
  
  #calculate LD between each SNP and the peak 2 SNP
  if(nrow(peak2) != 0){
    LDmatrix_peak2 <- genomat_peaks_i %>% dplyr::select(-c(SNP, compute_group, CHROM, POS, distance_peak1, peak1_SNP)) %>% 
      t() %>% cor()
    LDwithpeak2 <- LDmatrix_peak2[,paste0(peak2$SNP_ID)]
    genomat_peaks_i$LDpeak2 <- LDwithpeak2
    
    genomat_peaks_i <- genomat_peaks_i %>% mutate(distance_peak2 = POS - peak2$POS, peak2_SNP = peak2$SNP_ID)
  } else{genomat_peaks_i <- genomat_peaks_i %>% mutate(LDpeak2 = NA, distance_peak2 = NA, peak2_SNP = NA)
  } #end chromosome
  
  #join chromosome results with the others
 genomat_peaks_allchrom <- rbind(genomat_peaks_allchrom, genomat_peaks_i)
 
}

# Examine LD decay around top SNPs
LDdecay <- data.table::rbindlist(list(genomat_peaks_allchrom %>% dplyr::select(distance_peak1, LDpeak1),
                 genomat_peaks_allchrom %>% dplyr::select(distance_peak2, LDpeak2)), use.names=F) %>%
  filter(!is.na(distance_peak1)) %>%
  mutate(Rsquared = LDpeak1^2)

#Find average LD within 1 kb bins
bins=seq(1,max(LDdecay$distance_peak1, na.rm=T), by=1000)
my.means=rep(0, length(bins))
LD.averages=data.frame(bins, my.means)
for (i in 1:length(bins)) {
    data.interval=subset(LDdecay, (LDdecay$distance_peak1 >= bins[i] & LDdecay$distance_peak1 < (bins[i]+500)))
    LD.averages$my.means[i]=mean(data.interval$Rsquared)
}
plot(LD.averages$bins, LD.averages$my.means)
plot(LD.averages$bins, LD.averages$my.means, xlim=c(0, 80000))

# Visualize LD decay
ggplot(data= LDdecay, aes(x=abs(distance_peak1), y=Rsquared)) +
         geom_point() +
         geom_smooth(se = FALSE, method = "loess") +
         ggtitle("LD decay around peak SNPs") +
         xlab("distance (bp) to peak SNP") +
         ylab("correlation with peak SNP")

ggsave(filename=paste0(project_path, "figures/asreml_mvGWAS_LDdecay_unroundedgenos_weighted_",
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))


# Plot Manhattan colored by LD with peak SNP
all_SNP_results_peaks <- genomat_peaks_allchrom %>% dplyr::select(c(SNP, CHROM, POS, LDpeak1, distance_peak1, peak1_SNP, LDpeak2, distance_peak2, peak2_SNP)) %>% 
  mutate(CHROM = as.character(CHROM)) %>%
  mutate(distance_any = case_when(is.na(distance_peak2) | 
                                    abs(distance_peak1) < abs(distance_peak2) ~ distance_peak1, 
                                  TRUE ~ distance_peak2)) %>%
 mutate(Rsquared_any = case_when(is.na(distance_peak2) | 
                                   abs(distance_peak1) < abs(distance_peak2) ~ LDpeak1^2, 
                                 TRUE ~ LDpeak2^2)) %>%
  mutate(peak_SNP = case_when(SNP == peak1_SNP | SNP == peak2_SNP ~ "yes", T ~ "no")) %>%
  left_join(all_SNP_results_df, .) %>% 
  arrange(as.numeric(CHROM), as.numeric(POS))

#order factor levels so they display in order of chromosome
all_SNP_results_peaks$SNP <- factor(all_SNP_results_peaks$SNP, 
                                        levels = all_SNP_results_peaks$SNP)
all_SNP_results_peaks$CHROM <- as.factor(all_SNP_results_peaks$CHROM)
all_SNP_results_peaks$CHROM <- factor(all_SNP_results_peaks$CHROM, 
                                               levels = as.character(c(1:18)))

saveRDS(all_SNP_results_peaks, file=paste0(project_path, "output/all_SNP_results_MVWald_wPSY2KASPand5PC2_unrounded_weighted_annotations_effdir_", format(Sys.time(), "%m%d_%H%M"), ".RDS"))


## Plot Manhattan with joint p-values, colored by LD with peak SNP
ggplot(all_SNP_results_peaks, aes(x=POS, y=-log10(P), color=Rsquared_any)) +
  geom_point(aes(pch=peak_SNP, size=peak_SNP)) +
  scale_color_gradient2(low = "blue", mid="orange", high = "red", midpoint = 0.5, 
                          name= expression(textstyle("R"^2*" with starred SNP"))) +
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM, PSY2 cov, unrounded,\nweighted, with LD around peak SNPs")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.2) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.2, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed", lwd=0.2) #PSY2 causal SNP


ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_revised_unrounded_weighted_wLD_", format(Sys.time(), "%m%d_%H%M"), ".jpg"), width=12, height=5)


# Color Manhattan by effect direction same/different
colPalette2 <- c("#00BFC4","gray", "#F8766D")

ggplot(all_SNP_results_peaks, aes(x=POS, y=-log10(P), color=effectdir)) +
  geom_point(aes(pch=peak_SNP, size=peak_SNP)) +
  scale_color_manual(values = colPalette2, name="effect direction") +
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM, PSY2 covariate, unrounded,\nweighted, with effect direction")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.2) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.2, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
   geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed", lwd=0.2) #PSY2 causal SNP

ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_unrounded_weighted_wEffectDir_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))
```

Binomial test and Table 2
```{r}
all_SNP_results_peaks <- readRDS(file=paste0(project_path, "output/all_SNP_results_MVWald_wPSY2KASPand5PC2_unrounded_weighted_annotations_effdir_0717_1703.RDS"))


# Binomial test
peakSNPsonly <- all_SNP_results_peaks %>% filter(peak_SNP == "yes")
nrow(peakSNPsonly) #17, with 11 out of 17 in opposite directions
table(peakSNPsonly$effectdir)
binom.test(x=c(table(peakSNPsonly$effectdir)), p=0.5, alternative = "greater")
#p-value = 0.1662


# Reattach allele Ref/Alt information and MAF:
genomat_SNPs <- readRDS(file="./data/genomat_rev.RDS")
genomat_SNPinfo <- readRDS(file=paste0(project_path, "data/AllChrom_MergedCASSYellowDARTGenotypesSharedSites_REF19imputedAndFiltered_SNPinfo.RDS")) %>% filter(ID %in% colnames(genomat_SNPs)) %>% rename("ID" = "SNP_ID") 
genomat_SNPinfo$AF <- colSums(genomat_SNPs) / (nrow(genomat_SNPs)*2)


peakSNPsonly_alleles <- left_join(peakSNPsonly, genomat_SNPinfo %>% dplyr::select(SNP_ID,ALT,REF,AF)) %>%
  mutate(Minorallele = case_when(AF >= 0.5 ~ REF, AF < 0.5 ~ ALT),
         Majorallele = case_when(AF >= 0.5 ~ ALT, AF < 0.5 ~ REF)) %>%
  mutate(MAF = case_when(AF < 0.5 ~ AF, AF >= 0.5 ~ (1-AF) ))

table2 <- peakSNPsonly_alleles %>% arrange(P) %>% dplyr::select(SNP_ID,P,BetaSNP_TCICHK,BetaSNP_TCICHK.se, BetaSNP_DM,BetaSNP_DM.se,effectdir,REF,ALT,AF) 

# old method -- get total phenotypic variance for TCICHK and DM to calculate PVE
#totalv_TCICHK <- var(modeldata$TCICHK, na.rm=T) #0.4393742
#totalv_DM <- var(modeldata$DM, na.rm=T) #0.3920023

table2_rounded <- table2 %>%
  #calculate proportion variance explained (PVE)
  mutate(PVE_TCICHK = 100*(BetaSNP_TCICHK^2 * 2 * AF * (1-AF))/ nullmodel_TCICHKv,
         PVE_DM = 100*(BetaSNP_DM^2 * 2 * AF * (1-AF))/ nullmodel_DMv,
        PCE_cov =  100*(BetaSNP_TCICHK * BetaSNP_DM * 2 * AF * (1-AF))/ nullmodel_cov) %>%
  #round decimals for display in table
  mutate(across(c(BetaSNP_TCICHK:BetaSNP_DM.se, PVE_TCICHK:PCE_cov), ~ round(.x, 3)),
         AF = round(AF, 2)) %>%
  mutate(P = formatC(P, format = "e", digits = 2)) %>%
  mutate(TCeffectdir = case_when(sign(BetaSNP_TCICHK) == 1 ~ "+", sign(BetaSNP_TCICHK) == -1 ~ "-", T~NA),
         DMeffectdir = case_when(sign(BetaSNP_DM) == 1 ~ "+", sign(BetaSNP_DM) == -1 ~ "-", T~NA)) %>%
  mutate(EffectDirection = paste0(TCeffectdir, "/", DMeffectdir)) %>%
  relocate(EffectDirection, .after=BetaSNP_DM.se) %>%
  mutate(pvalue = case_when(as.numeric(P) < Gao.cutoff ~ paste0(P, " **"), 
                             T ~ paste0(P, " *"))) %>%
  relocate(pvalue, .after=SNP_ID)  %>%
  dplyr::select(-c(TCeffectdir, DMeffectdir, effectdir, P))

## TABLE 2
View(table2_rounded)

write.csv(table2_rounded, file=paste0(project_path, "output/table2_wPSY2KASPand5PCs_revised_unroundedgenotypes_weighted_", format(Sys.time(), "%m%d_%H%M"), ".csv"))


#total PCE by all peak SNPs:
sum(table2_rounded$PCE_cov)

# PCE by PSY2 SNP:
PSY2effects <- as.data.frame(summary(nullmodel_PSY2, coef=T)$coef.fixed)[c("trait_TCICHK:S1_24155522","trait_DM:S1_24155522"),1]
PSY2AF <- sum(BLUPs_PSY2_data$S1_24155522)/(nrow(BLUPs_PSY2_data))

PCE_PSY2 <- 100*((PSY2effects[1] * PSY2effects[2] * 2 * PSY2AF * (1-PSY2AF))/ nullmodel_cov)
print(PCE_PSY2) #5.57%

# PVE by  PSY2 SNP:
PVE_TCICHK_PSY2 <- 100*(PSY2effects[1]^2 * 2 * PSY2AF * (1-PSY2AF))/ nullmodel_TCICHKv
print(PVE_TCICHK_PSY2) #2.44%

PVE_DM_PSY2 <-  100*(PSY2effects[2]^2 * 2 * PSY2AF * (1-PSY2AF))/ nullmodel_DMv
print(PVE_DM_PSY2) #5.01%
```

Test significance of negative genetic covariance in model with PSY2 and all SNPs passing the relaxed threshold:
```{r}
PeakSNPsdata <- genomat_t %>%
  filter(genomat_t$SNP %in% peakSNPsonly$SNP_ID) 

rownames(PeakSNPsdata) <- PeakSNPsdata$SNP
PeakSNPsdata_t <- as.data.frame(t(PeakSNPsdata %>% dplyr::select(-c(SNP, compute_group))))
colnames(PeakSNPsdata_t) <- paste0(PeakSNPsdata$SNP)
PeakSNPsdata_t$germplasmName <- rownames(PeakSNPsdata_t)

modeldata <- left_join(BLUPs_PSY2_data, PeakSNPsdata_t, by="germplasmName") %>% 
      mutate(germplasmName = as.factor(germplasmName))
    
    
#fit model with all suggestively associated SNPs and PSY2 effect
model_allpeakSNPs <- asreml(data= modeldata, 
       fixed = formula (paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + ", 
                              paste0("PC", 1:5, collapse="+"), "+", 
                              paste0("trait:", peakSNPsonly$SNP_ID, collapse="+")) ),
       random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
       residual =  ~units:us(trait), weights=acc_weight, maxit=20)

saveRDS(model_allpeakSNPs, file=paste0(project_path, "output/fullmodel_PSY2KASP_revised_unrounded_weighted_15peakSNPs.RDS"))


#fix residual genetic variance parameters to 0
model_allpeakSNPs_copy <- model_allpeakSNPs
G_constrained <- initG
names(G_constrained) <- c("F", "U", "U")
G_constrained[1] <- 0
E_constrained <- model_allpeakSNPs_copy$vparameters[5:7]
names(E_constrained) <- c("F", "U", "U")
E_constrained[2] <- 0


model_allpeakSNPs_reducedmanually_startvals <- asreml(data= modeldata, 
    fixed = formula (paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + ", 
                              paste0("PC", 1:5, collapse="+"), "+", 
                              paste0("trait:", peakSNPsonly$SNP_ID, collapse="+")) ),
  random = ~ corgh(trait, init=G_constrained):vm(germplasmName, Ginverse),
  residual = ~ units:us(trait, init=E_constrained),  weights = acc_weight, 
  maxit = 20, start.values=TRUE)
model_allpeakSNPs_reducedmanually_startvals$R.param$`units:trait`$trait$con <- c("P","F", "P")

model_allpeakSNPs_reducedmanually <- asreml(data= modeldata, 
    fixed = formula (paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + ", 
                              paste0("PC", 1:5, collapse="+"), "+", 
                              paste0("trait:", peakSNPsonly$SNP_ID, collapse="+")) ),
  random = ~ corgh(trait, init=G_constrained):vm(germplasmName, Ginverse),
  residual = ~ units:us(trait, init=E_constrained),  weights = acc_weight, 
  maxit = 20, R.param=model_allpeakSNPs_reducedmanually_startvals$R.param)


#fit reduced model with the 'residual' polygenic genetic covariance set to 0
# model_allpeakSNPs_0cov <- asreml(data = modeldata, 
#        fixed= formula (paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + ", 
#                               paste0("PC", 1:5, collapse="+"), "+", 
#                               paste0("trait:", peakSNPsonly$SNP_ID, collapse="+"))),
#        random = ~ diag(trait):vm(germplasmName, Ginverse),
#        residual =  ~units:diag(trait), weights=acc_weight, maxit=20)

LRTresults <- lrt.asreml(model_allpeakSNPs_reducedmanually, model_allpeakSNPs, boundary=T)
print(LRTresults)
saveRDS(LRTresults, file=paste0(project_path, "output/LRT_residual_gen_covar_revised_unrounded_weighted.RDS"))



#LD between top SNPs:
peakSNPscormat <- cor( modeldata %>% dplyr::select(S1_24155522, peakSNPsonly$SNP_ID) )
peakSNPs_Rsquared <- peakSNPscormat^2
peakSNPs_Rsquared_PSY2 <- ((as.data.frame(peakSNPscormat)$S1_24155522)[-1])^2
max(peakSNPs_Rsquared_PSY2)
```
A significant likelihood ratio test here (0.002498) indicates that the model with negative residual genetic covariance is a significantly better fit than one without it. We can conclude that there is significant negative residual genetic covariance even after accounting for all of the suggestively associated SNPs.



Zoom in on chromosome 1: 
```{r, eval=F}
#do LD calculation for all significant SNPs on chromosome 1:
genomat_ROG <- genomat_t %>%
  mutate(CHROM= as.integer(gsub("S", "", SNP) %>% gsub("_[0-9].*", "", .)),
         POS= as.numeric(gsub("^S.*_", "", SNP)))

genomat_peaks_1 <- genomat_ROG %>% filter(CHROM == "1")
rownames(genomat_peaks_1) <- genomat_peaks_1$SNP
  
# pull out the names of any SNPs above the Gao cutoff
topSNPs_chri <- topSNPs %>% filter(CHROM == "1") %>% arrange(P)
num_sig_snps <- seq(1, sum(topSNPs_chri$P < Gao.cutoff))
for(i in 1:length(num_sig_snps)){
  assign ( paste0("peak", i), topSNPs_chri[num_sig_snps[i],] )
}
peak2 <- topSNPs_chri[2,]#manually add the second most significant SNP which is below Gao cutoff

#calculate LD between each SNP and the peak SNPs
snp_cor_mat <- genomat_peaks_1 %>% dplyr::select(-c(SNP, compute_group, CHROM, POS)) %>% 
    t() %>% cor()

LDwithpeak1 <- snp_cor_mat[,paste0(peak1$SNP_ID)]
LDwithpeak2 <- snp_cor_mat[,paste0(peak2$SNP_ID)]

genomat_peaks_1 <- genomat_peaks_1 %>% mutate(distance_peak1 = POS - as.numeric(peak1$POS), 
                                              peak1_SNP = peak1$SNP_ID,
                                              LDpeak1 = LDwithpeak1,
                                              distance_peak2 = POS - as.numeric(peak2$POS),
                                              peak2_SNP = peak2$SNP_ID,
                                              LDpeak2 = LDwithpeak2,
                                              Rsquared_peak1 = LDwithpeak1^2,
                                              Rsquared_peak2 = LDwithpeak2^2)

#Prepare to plot
# Plot Manhattan colored by LD with peak SNP
chr1_results_peaks <- genomat_peaks_1 %>% mutate(CHROM = as.character(CHROM)) %>%
  mutate(isthispeak1 = case_when(SNP == peak1_SNP  ~ "yes", T ~ "no")) %>%
  mutate(isthispeak2 = case_when(SNP == peak2_SNP  ~ "yes", T ~ "no")) %>%
  left_join(all_SNP_results_df %>% filter(CHROM=="1"), .) %>% 
  arrange(as.numeric(CHROM), as.numeric(POS)) %>%
  filter(!duplicated(SNP_ID)) %>%
  mutate(SNP = factor(SNP, levels = SNP))

saveRDS(chr1_results_peaks, file=paste0(project_path, "output/chr1_results_wPSY2KASPand5PCs_revised_weighted_wpeak_annotations.RDS"))
```

```{r}
chr1_results_peaks <- readRDS(paste0(project_path,"output/chr1_results_wPSY2KASPand5PCs_revised_weighted_wpeak_annotations.RDS"))

## Plot chr 1 Manhattan with joint p-values, colored by LD with peak SNP
ggplot(chr1_results_peaks %>% filter(POS > 20000000), aes(x=POS, y=-log10(P), color=Rsquared_peak1)) +
  geom_point(aes(pch=isthispeak1, size=isthispeak1)) +
  scale_color_gradient2(low = "blue", mid="orange", high = "red", midpoint = 0.5, name=
                         expression(textstyle("R"^2*" with starred SNP"))) +
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chr. 1 physical position (bp)") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate and PCs, weighted")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  theme(strip.background = element_rect(fill = "white")) +
  geom_vline(aes(xintercept=24155522), color="orange", lty="dashed") + #causal PSY2 SNP 
  annotate("segment", x = c(19400000, 19600000), xend = c(19600000, 19800000), y = c(-0.1, -0.1), yend = c(0.1, 0.1)) +
  coord_cartesian(clip = "off", xlim = c(19000000, 35000000)) +
  #annotate("rect", xmin = 24133741, xmax = 24170185, ymin = 0, ymax = 6, alpha = 0.2, fill="orange") #PSY2 haplotype region
  annotate("rect", xmin = 24153419, xmax = 24156720, ymin = 0, ymax = 6, alpha = 0.2, fill="orange") #PSY2 coding region


ggsave(filename=paste0(project_path, "figures/mvGWAS_wPSY2KASPandPCs_weighted_Chr1Manhattan_wLD_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))

ggsave(filename=paste0(project_path, "figures/mvGWAS_wPSY2KASPandPCs_weighted_Chr1Manhattan_wLD_", 
                       format(Sys.time(), "%m%d_%H%M"), "_wide.jpg"), height=5, width=8, units="in")


## Plot chr 1 Manhattan with joint p-values, colored by LD with peak 2 SNP
ggplot(chr1_results_peaks, aes(x=POS, y=-log10(P), color=Rsquared_peak1)) +
  geom_point(aes(pch=isthispeak1, size=isthispeak1)) +
  scale_color_gradient2(low = "blue", mid="orange", high = "red", midpoint = 0.5, name=
                           expression(textstyle("R"^2*" with starred SNP"))) +
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chr. 1 physical position (bp)") +
  ggtitle(paste0("MLMM GWAS for TC and DM, weighted")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  theme(strip.background = element_rect(fill = "white"))  +
  geom_vline(aes(xintercept=24155522), color="orange", lty="dashed")

ggsave(filename=paste0(project_path, "figures/mvGWAS_wPSY2KASPandPCs_weighted_Chr1Manhattan_wLD_pk2_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))

ggsave(filename=paste0(project_path, "figures/mvGWAS_wPSY2KASPandPCs_weighted_Chr1Manhattan_wLD_pk2_", 
                       format(Sys.time(), "%m%d_%H%M"), "_wide.jpg"), height=4, width=6, units="in")


```


Test LD of chr 1 peak SNP with PSY2:
```{r}
genomat_chr1qtl <- genomat_plusPSY %>% filter(SNP == "S1_24155522" | SNP == "S1_23484417") %>%
  column_to_rownames("SNP") %>%
  dplyr::select(-c(CHROM, POS))

LDchr1 <- cor(t(genomat_chr1qtl))
(LDchr1[1,2])^2
```


Test LD of all bivariate peak SNPs with each other:
```{r}
topSNPs_LD <- (cor(t(genomat_topSNPs)))^2
diag(topSNPs_LD) <- NA

topSNPs_LD_vector <- topSNPs_LD[lower.tri(topSNPs_LD, diag=F)]
topSNPs_LD_vector[which.max(topSNPs_LD_vector)] # 0.095
```




# Model comparison --
```{r}
# Compare association models: bivariate (1), univariate with other trait covariate (2 & 3), and univariate (4a & 4b)
 
#get top SNP info
alltopSNPlist <- peakSNPsonly$SNP
topSNPs <- all_SNP_results_MVWald %>% filter(SNP_ID %in% alltopSNPlist)
genomat_topSNPs <- genomat_ROG %>% filter(SNP %in% alltopSNPlist) %>%
  dplyr::select(-c(SNP, compute_group, CHROM, POS))


# Loop across all top SNPs 
# initiate storage
model_storage <- list()
model_comparison_allSNPs <- list()
  
for(i in 1:nrow(genomat_topSNPs)){
  print(i)
    
  SNPi <- rownames(genomat_topSNPs)[i]
  SNPidata <- as.data.frame(t(genomat_topSNPs[i, ]))
  colnames(SNPidata) <- paste0(SNPi)
  SNPidata$germplasmName <- rownames(SNPidata)
  
  modeldata <- left_join(BLUPs_PSY2_data, SNPidata, by="germplasmName") %>% 
      mutate(germplasmName = as.factor(germplasmName)) %>%
      filter(!is.na(TCICHK) & !is.na(DM))

  
  #model 1: SNP is jointly associated with bivariate phenotype
  SNPimodel1 <- asreml(data = modeldata,
        fixed = formula(paste0("cbind(TCICHK, DM) ~ trait + trait:", SNPi, 
                               " + trait:S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
        residual = ~units:us(trait), weights=acc_weight, maxit=30, trace=F)
  
   #predict SNP effects
    SNPlevel <- list(1)
    names(SNPlevel) <- SNPi
    predictSNPeffects1 <- predict.asreml(SNPimodel1, vcov=T, 
                                        classify=paste0("trait:", SNPi), 
                                        only=paste0("trait:", SNPi), 
                                        levels=SNPlevel)
    
    Betas_i <- pivot_wider(predictSNPeffects1$pvals, names_from = "trait", 
                           values_from=c("predicted.value", "std.error"))
    Cov_i <- predictSNPeffects1$vcov
    
    #Bivariate Wald test:
    WaldTest1bv <- aod::wald.test(Sigma=predictSNPeffects1$vcov,
                               b=predictSNPeffects1$pvals$predicted.value, Terms=1:2)
    
  
  WaldTest1 <- as.data.frame(wald.asreml(SNPimodel1))
  WaldTest1$AIC <- summary(SNPimodel1)$aic
  WaldTest1$BIC <- summary(SNPimodel1)$bic
  WaldTest1$bvP <- WaldTest1bv$result$chi2["P"]
  WaldTest1$loglik <- summary(SNPimodel1)$loglik

  
  #model 2: SNP is associated with TC while accounting for DM
  SNPimodel2 <- asreml(data = modeldata,
        fixed = formula(paste0("TCICHK ~ 1 + DM*", SNPi, " + S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ vm(germplasmName, Ginverse),
        residual = ~units, weights=TCICHK_weight, maxit=20, trace=F)
  
  WaldTest2 <- as.data.frame(wald.asreml(SNPimodel2))
  WaldTest2$AIC <- summary(SNPimodel2)$aic
  WaldTest2$BIC <- summary(SNPimodel2)$bic
  WaldTest2$loglik <- summary(SNPimodel2)$loglik
    
  
  #model 3: SNP is associated with DM while accounting for TC
  SNPimodel3 <- asreml(data = modeldata,
        fixed = formula(paste0("DM ~ 1 + TCICHK*", SNPi, " + S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ vm(germplasmName, Ginverse),
        residual = ~units, weights=DM_weight, maxit=20, trace=F)
  
  WaldTest3 <- as.data.frame(wald.asreml(SNPimodel3))
  WaldTest3$AIC <- summary(SNPimodel3)$aic
  WaldTest3$BIC <- summary(SNPimodel3)$bic
  WaldTest3$loglik <- summary(SNPimodel3)$loglik

  
  #model 4a: SNP is associated with TC (univariate)
  SNPimodel4a <- asreml(data = modeldata,
        fixed = formula(paste0("TCICHK ~ 1 + ", SNPi, "+ S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ vm(germplasmName, Ginverse),
        residual = ~units, weights=TCICHK_weight, maxit=20, trace=F)
  WaldTest4a <- as.data.frame(wald.asreml(SNPimodel4a))
  WaldTest4a$AIC <- summary(SNPimodel4a)$aic
  WaldTest4a$BIC <- summary(SNPimodel4a)$bic
  WaldTest4a$loglik <- summary(SNPimodel4a)$loglik

  
  #model 4b: SNP is associated with DM (univariate)
  SNPimodel4b <- asreml(data = modeldata,
        fixed = formula(paste0("DM ~ 1 + ", SNPi, " + S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ vm(germplasmName, Ginverse),
        residual = ~units, weights = DM_weight, maxit=20, trace=F)
  WaldTest4b <- as.data.frame(wald.asreml(SNPimodel4b))
  WaldTest4b$AIC <- summary(SNPimodel4b)$aic
  WaldTest4b$BIC <- summary(SNPimodel4b)$bic
  WaldTest4b$loglik <- summary(SNPimodel4b)$loglik

  # Save results from this SNP:
   model_stats_storage_i <- list(WaldTest1, WaldTest2, WaldTest3, WaldTest4a, WaldTest4b)
     
   model_comparison_allSNPs[[i]] <- model_stats_storage_i

} #end SNP
  

saveRDS(model_comparison_allSNPs, file=paste0(project_path, 
                  "output/mvGWAS_asreml_output/topSNP_model_comparison_Waldstats_wPSY2KASP_5PCs_revised_weighted_", format(Sys.time(), "%m%d"), ".RDS"))
```

Evaluate model comparison for peak SNPs
```{r}
lastrundate <- "0720"
model_comparison_allSNPs <- readRDS(file=paste0(project_path,
                  "output/mvGWAS_asreml_output/topSNP_model_comparison_Waldstats_wPSY2KASP_5PCs_revised_weighted_", lastrundate, ".RDS"))

#model 1: bivariate; model 2: TCICHK with DM covariate, model 3: TCICHK with DM covariate, model 4a: univariate TCICHK, model 4b: univariate DM

modelcomp_summary_table <- data.frame()
for(i in 1:nrow(genomat_topSNPs)){
  SNPname <- rownames(genomat_topSNPs)[i]

  mod1 <- model_comparison_allSNPs[[i]][[1]][paste0("trait:",SNPname),c("Pr(Chisq)", "AIC", "BIC", "loglik")]
  colnames(mod1) <- paste0("1_", colnames(mod1))
  
  mod2 <- model_comparison_allSNPs[[i]][[2]][SNPname,c("Pr(Chisq)", "AIC", "BIC", "loglik")]
  colnames(mod2) <- paste0("2_", colnames(mod2))
  
  mod3 <- model_comparison_allSNPs[[i]][[3]][SNPname,c("Pr(Chisq)", "AIC", "BIC", "loglik")]
  colnames(mod3) <- paste0("3_", colnames(mod3))
  
  mod4a <- model_comparison_allSNPs[[i]][[4]][SNPname,c("Pr(Chisq)", "AIC", "BIC", "loglik")]
  colnames(mod4a) <- paste0("4a_", colnames(mod4a))
  
  mod4b <- model_comparison_allSNPs[[i]][[5]][SNPname,c("Pr(Chisq)", "AIC", "BIC", "loglik")]
  colnames(mod4b) <- paste0("4b_", colnames(mod4b))
  

AICcomp <- c(mod1$`1_AIC`,mod2$`2_AIC`,mod3$`3_AIC`,mod4a$`4a_AIC`, mod4b$`4b_AIC`)
#BICcomp <- c(mod1$`1_BIC`,mod2$`2_BIC`,mod3$`3_BIC`,mod4a$`4a_BIC`, mod4b$`4b_BIC`)
Pcomp <- c(mod1$`1_Pr(Chisq)`,mod2$`2_Pr(Chisq)`,mod3$`3_Pr(Chisq)`) #mod4a$`4a_Pr(Chisq)`, mod4b$`4b_Pr(Chisq)`)
loglikcomp <- c(mod1$`1_loglik`, mod2$`2_loglik`, mod3$`3_loglik`, mod4a$`4a_loglik`, mod4b$`4b_loglik`)


tosave <- cbind(mod1, mod2, mod3, mod4a, mod4b)
tosave$lowestAICmod <- which.min(AICcomp)
#tosave$lowestBICmod <- which.min(BICcomp)
tosave$lowestP <- which.min(Pcomp)
tosave$highestloglik <- which.max(loglikcomp)

rownames(tosave) <- gsub("trait:", "", rownames(tosave))
modelcomp_summary_table <- rbind(modelcomp_summary_table, tosave)
}

View(modelcomp_summary_table)

#format for table 3:
table3 <- data.frame(PeakSNP = gsub("S","",rownames(modelcomp_summary_table)), 
           BivariateAIC = round(modelcomp_summary_table$`1_AIC`, 2),
           TC_AIC = round(modelcomp_summary_table$`2_AIC`, 2),
           DM_AIC = round(modelcomp_summary_table$`3_AIC`, 2),
           #Bivariateloglik = modelcomp_summary_table$`1_loglik`, 
           #TC_loglik = modelcomp_summary_table$`2_loglik`, 
           #DM_loglik = modelcomp_summary_table$`3_loglik`,
           BivariateP = modelcomp_summary_table$`1_Pr(Chisq)`,
           TC_P = modelcomp_summary_table$`2_Pr(Chisq)`,
           DM_P = modelcomp_summary_table$`3_Pr(Chisq)`)


table3_formatted <- table3 %>% arrange(BivariateP) %>%
  mutate(BivariateP = case_when(as.numeric(BivariateP) < Gao.cutoff ~ 
                                  paste0(formatC(BivariateP, format="e", digits=2), " **"),
                                as.numeric(BivariateP) < relaxed.cutoff ~ 
                                  paste0(formatC(BivariateP, format="e", digits=2),  " *"),
                             T ~ paste0(formatC(BivariateP, format="e", digits=2))) ) %>%
  mutate(TC_P = case_when(as.numeric(TC_P) < Gao.cutoff ~ 
                                  paste0(formatC(TC_P, format="e", digits=2), " **"),
                                as.numeric(TC_P) < relaxed.cutoff ~ 
                                  paste0(formatC(TC_P, format="e", digits=2),  " *"),
                             T ~ paste0(formatC(TC_P, format="e", digits=2))) ) %>%
  mutate(DM_P = case_when(as.numeric(DM_P) < Gao.cutoff ~ 
                                  paste0(formatC(DM_P, format="e", digits=2), " **"),
                                as.numeric(DM_P) < relaxed.cutoff ~ 
                                  paste0(formatC(DM_P, format="e", digits=2),  " *"),
                             T ~ paste0(formatC(DM_P, format="e", digits=2))) )

View(table3_formatted)


table3 %>% rmarkdown::paged_table()
write.csv(table3_formatted, file="/workdir/ssv42/mvGWAS/output/table3_modelcomparison_wPSY2SNP_5PCs_revised_weighted.csv")

```
Note: I am using the bivariate p-value from the original GWAS model, which is slightly different than the one listed here, because this analysis required me to subset the data to remove any observations with missing DM or TCICHK data, which are necessary for the respective univariate models but not needed in the bivariate model.






Get effect estimates using mixed model fit on untransformed phenotype BLUPs
```{r}
#Get drgBLUPs from untransformed phenotypes:
OGBLUPs <- readRDS(file=paste0(project_path, "data/untransformed_OGphenos_TCICHK_DM_BLUPs.RDS"))

# join OGpheno BLUPs with rest of data
OGBLUPs_PSY2_data <- left_join(BLUPs_PSY2_data, OGBLUPs, by=c("germplasmName")) 



##Re-run bivariate mixed model with OGBLUPs for each of the top SNPs
genomat_topSNPs <- genomat_SNPs %>% filter(SNP %in% topSNPs$SNP_ID) %>%
  dplyr::select(-c(SNP, compute_group, CHROM, POS))

OGBLUPs_topSNP_results_df <- data.frame()

for(i in 1:nrow(genomat_topSNPs)){
 
  SNPi <- rownames(genomat_topSNPs)[i];  print(SNPi)
  SNPidata <- as.data.frame(t(genomat_topSNPs[i,]))
  colnames(SNPidata) <- paste0(SNPi)
  SNPidata$germplasmName <- rownames(SNPidata)
  
  modeldata <- left_join(OGBLUPs_PSY2_data, SNPidata, by=c("germplasmName")) %>%
    mutate(germplasmName = as.factor(germplasmName))

OGpheno_biv_model <- asreml(data = modeldata,
        fixed = formula(paste0("cbind(TCICHK_OG, DM_OG) ~ trait + trait:", SNPi, 
                               " + trait:S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
        residual = ~units:us(trait), weights=acc_weight, maxit=30, trace=F)
  
   #predict SNP effects
    SNPlevel <- list(1)
    names(SNPlevel) <- SNPi
    predictSNPeffects1 <- predict.asreml(OGpheno_biv_model, vcov=T, 
                                        classify=paste0("trait:", SNPi), 
                                        only=paste0("trait:", SNPi), 
                                        levels=SNPlevel)
    
    Betas_i <- pivot_wider(predictSNPeffects1$pvals, names_from = "trait", 
                           values_from=c("predicted.value", "std.error"))
    Cov_i <- predictSNPeffects1$vcov
    
    #Bivariate Wald test:
    WaldTest1bv <- aod::wald.test(Sigma=predictSNPeffects1$vcov,
                               b=predictSNPeffects1$pvals$predicted.value, Terms=1:2)
    
    #Store results
    #Save results
    OGBLUPs_topSNP_results_df[i,1:10] <- cbind(Betas_i, Cov_i[1,2], t(WaldTest1bv$result$chi))

    rownames(OGBLUPs_topSNP_results_df)[i] <- colnames(Betas_i)[1]
    
    
}
OGBLUPs_topSNP_results_df$SNP_ID <- rownames(OGBLUPs_topSNP_results_df)


  
saveRDS(OGBLUPs_topSNP_results_df, file=paste0(project_path, 
          "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_betaestimates_untransformedphenos_revised_unrounded_weighted_",  format(Sys.time(), "%m%d"), ".RDS"))



table2_expanded <- left_join(table2_rounded, OGBLUPs_topSNP_results_df, by=c("SNP_ID")) 

betaSNP_comparison <- table2_expanded %>% dplyr::select(SNP_ID, BetaSNP_TCICHK, predicted.value_TCICHK_OG, BetaSNP_DM, predicted.value_DM_OG, pvalue, P)
colnames(betaSNP_comparison) <- c("SNP","BoxCox_Beta_TCICHK","Original_Beta_TCICHK","BoxCox_Beta_DM","Original_Beta_DM", "P_trans", "P_OG")


View(betaSNP_comparison)

# plot(betaSNP_comparison$BoxCox_Beta_TCICHK, betaSNP_comparison$Original_Beta_TCICHK,
#      xlab="Box-Cox transformed SNP effect estimate", 
#      ylab="Untransformed data SNP effect estimate", 
#      main="TCICHK")
# 
# plot(betaSNP_comparison$BoxCox_Beta_DM, betaSNP_comparison$Original_Beta_DM,
#      xlab="Box-Cox transformed SNP effect estimate", 
#      ylab="Untransformed data SNP effect estimate", 
#      main="DM")

# Compare effect estimates from Box-Cox transformed data to estimates from untransformed data
TCBoxCoxmod <- lm(betaSNP_comparison$Original_Beta_TCICHK ~ betaSNP_comparison$BoxCox_Beta_TCICHK)
TCBoxCoxmod$coefficients
summary(TCBoxCoxmod)$r.squared

DMBoxCoxmod <- lm(betaSNP_comparison$Original_Beta_DM ~ betaSNP_comparison$BoxCox_Beta_DM)
DMBoxCoxmod$coefficients
summary(DMBoxCoxmod)$r.squared


ggplot(betaSNP_comparison, aes(BoxCox_Beta_TCICHK, Original_Beta_TCICHK)) +
  geom_point(aes(color="TCICHK")) +
  geom_point(aes(BoxCox_Beta_DM, Original_Beta_DM, color="DM")) +
  geom_abline(intercept= TCBoxCoxmod$coefficients[1], slope=TCBoxCoxmod$coefficients[2], color ="#F5793A") +
  geom_abline(intercept= DMBoxCoxmod$coefficients[1], slope=DMBoxCoxmod$coefficients[2], color ="#0F2080") +
  xlab("SNP effect estimate from Box-Cox transformed data") +
  ylab("SNP effect estimate from untransformed data") +
  scale_color_manual(values = c("TCICHK" = "#F5793A", "DM" = "#0F2080"),
                     labels = c( expression("DM (%)"), expression("TCICHK ("*mu*"g/g) fresh weight") )) +
  guides(color = guide_legend("Trait")) +
  theme(legend.text = element_text(hjust = 0))
  

ggsave(file=paste0(project_path, "figures/BoxCox_untransformed_effect_estimate_revised_weighted_comparison.png"))
  
  

range(betaSNP_comparison$Original_Beta_TCICHK)
range(betaSNP_comparison$Original_Beta_DM)


#Effect estimates of top SNPs on regular scale:
betaSNP_comparison %>% filter(SNP =="S1_23484417")
betaSNP_comparison %>% filter(SNP == "S10_2963170")
```

Allele frequency and effect sizes:
Boxplots:
```{r}
#Read in original (transformed) phenotype data -- just take mean for each accession
OGPhenos <- readRDS(file=paste0(project_path, "data/trialdata_subset_CASS_TCICHK_H2filterTRUE.RDS")) %>%
  group_by(germplasmName) %>%
  summarise(meanTCICHK = mean(TCICHK, na.rm=T), meanDM = mean(DM, na.rm=T)) %>%
  filter(!is.na(meanTCICHK))

# Plot boxplots to visualize peak SNP effects on raw phenotypes
GaopeakSNPs <- c("S1_24155522", "S1_23484417", "S10_2963170")
GaopeakSNPsdata <- as.data.frame(t(genomat_topSNPs[GaopeakSNPs,]))
GaopeakSNPsdata$germplasmName <- rownames(GaopeakSNPsdata)
  
boxplotdata <- left_join(OGBLUPs_PSY2_data, GaopeakSNPsdata, by=c("germplasmName")) %>%
  left_join(OGPhenos, by="germplasmName") %>%
  mutate(germplasmName = as.factor(germplasmName))
  
for(i in 1:length(GaopeakSNPs)){
  SNPi <- GaopeakSNPs[i]
  
  CombinedBoxplot_i <- ggplot(boxplotdata) + 
    geom_boxplot(aes(x=as.factor(round(.data[[SNPi]])), y=meanTCICHK), color="orange", alpha=0.7) +
    geom_boxplot(aes(x=as.factor(round(.data[[SNPi]])), y=meanDM), color="blue", alpha=0.7) +
      xlab(paste0(SNPi)) +
      ggtitle(paste0(SNPi, " effects on TCICHK and DM")) +
      #ylab("mean TCICHIK (ug/g) / mean DM (%)") +
      theme(axis.title.y = element_blank(),
            plot.margin = margin(10, 10, 10, 70)) +
      coord_cartesian(clip = 'off') +
      annotate("text", x = 0, y = 44, label = "DM (%)", 
             color = "blue", angle = 90, hjust = 1.5, vjust=-2.5, size = 5) +
      annotate("text", x = 0, y = 35, label = bquote("TCICHK (" * mu * "g/g)"), 
             color = "orange", angle = 90, hjust = 1.5, vjust=-1.8, size = 5)
  
  ggsave(CombinedBoxplot_i, file=paste0(project_path, "figures/boxplot_revised_roundedgenos_", SNPi, ".png"), height=5, width=7, units="in")
  
  print(CombinedBoxplot_i)
  
  
  dotplot_i <- ggplot(boxplotdata, aes(x=TCICHK, y=DM, color=as.factor(round(.data[[SNPi]])))) +
    geom_point() +
    scale_color_discrete( name = SNPi )
  
   ggsave(dotplot_i, file=paste0(project_path, "figures/dotplot_revised_roundedgenos_", SNPi, ".png"), height=5, width=7, units="in")
   print(dotplot_i)
}


ggplot(boxplotdata, aes(x=TCICHK, y=DM, color=as.factor(round(S10_2963170)))) +
    geom_point() +
    scale_color_discrete( name = "S10_2963170", labels = c("AA", "AC", "CC"))
  
ggsave(file=paste0(project_path, "figures/dotplot_revised_roundedgenos_allelelabeled_S10_2963170.png"), height=5, width=7, units="in")


ggplot(boxplotdata, aes(x=TCICHK, y=DM, color=as.factor(round(S1_23484417)))) +
    geom_point() +
    scale_color_discrete( name = "S1_23484417", labels = c("CC", "CG", "GG"))
  
ggsave(file=paste0(project_path, "figures/dotplot_revised_roundedgenos_allelelabeled_S1_23484417.png"), height=5, width=7, units="in")


ggplot(boxplotdata, aes(x=TCICHK, y=DM, color=as.factor(round(S1_24155522)))) +
    geom_point() +
    scale_color_discrete( name = "S1_24155522", labels = c("CC", "CA", "AA"))
  
ggsave(file=paste0(project_path, "figures/dotplot_revised_roundedgenos_allelelabeled_S1_24155522.png"), height=5, width=7, units="in")


```
































#########################
OLD code below:
#########################

## Examine the unweighted results
Make Q-Q plots
```{r}
all_SNP_results_df <- readRDS(file=paste0(project_path,
            "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_allSNPresults_0716.RDS"))

all_SNP_results_MVWald <- all_SNP_results_df

png(paste0(project_path, "figures/mvGWAS_wPSY2KASPand5PCs_unroundedgenos_QQplot_", dataset, "_",
                    format(Sys.time(), "%m%d"), ".png"), 
        width=6, height=6, res=250, units="in")
 
qqman::qq(all_SNP_results_MVWald$P, main="Q-Q plot of mvGWAS with reimputed genotypes")

dev.off()

# check the distribution of P-values
ggplot(all_SNP_results_MVWald, aes( BetaSNP_TCICHK/BetaSNP_TCICHK.se, 
                                    BetaSNP_DM/BetaSNP_DM.se, 
                                    color=-log10(P))) +
  ggtitle("w PSY2 covariate and PCs") +
  geom_point()


# check the distribution of the polygenic covariance between TC and DM -- is it still negative?
hist(all_SNP_results_MVWald$resid_gen_cov, main="residual genetic covariance between TC/DM \nafter accounting for PSY2 and SNP", xlab="covariance estimate")

range(all_SNP_results_MVWald$resid_gen_cov, na.rm=T)

```


Make Manhattan plot 
```{r}
all_SNP_results_MVWald$CHROM <- gsub("_[0-9]*","",all_SNP_results_MVWald$SNP_ID) %>% gsub("S","",.)
all_SNP_results_Bivar <- all_SNP_results_MVWald

# Set significance cut-off: Gao-modified Bonferroni
Gao.cutoff <- 1.367e-05

all_SNP_results_Bivar$POS <- as.numeric(all_SNP_results_Bivar$POS)
MVresults_sort <- all_SNP_results_Bivar %>% arrange(as.numeric(CHROM), as.numeric(POS))

#order factor levels so they display in order of chromosome
MVresults_sort$SNP_ID <- factor(MVresults_sort$SNP_ID, levels = MVresults_sort$SNP_ID)
MVresults_sort$CHROM <- as.factor(MVresults_sort$CHROM)
MVresults_sort$CHROM <- factor(MVresults_sort$CHROM, levels = as.character(c(1:18)))


#set color scheme
colPalette <- rep(c("blue4","pink3"),9)

## Plot Manhattan with bivariate Wald test p-values
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color=CHROM), size=0.5) +
  scale_color_manual(values=colPalette) +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate and unrounded genotypes")) +
  guides(color="none")+
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(legend.position = "none",
          strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP
#PSY2 v6 is 24153419 - 24156720

ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_unrounded_", format(Sys.time(), "%m%d_%H%M"), ".jpg"), width=12, height=5)


#for poster:
ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_unrounded_", dataset, "_", format(Sys.time(), "%m%d_%H%M"), "_wide.jpg"), width=12, height=3)


saveRDS(MVresults_sort, file=paste0(project_path, "output/MVresults_sorted_PSY2KASP_5PCs_revision_unroundedgenotypes.RDS"))

```

Color Manhattan plot by SNP covariance
```{r}
hist(MVresults_sort$BetaSNP_TCICHKDM.cov, main="Distribution of SNP TCICHK/DM effects covariance", xlab="SNP effect covariance")
hist((MVresults_sort %>% filter(P < 0.001))$BetaSNP_TCICHKDM.cov, main="Distribution of top SNP TCICHK/DM effects covariance for top SNPs", xlab="SNP effect covariance")


MVresults_sort$effectdir <- case_when(
      MVresults_sort$P > 0.001 ~ "NA",
      MVresults_sort$BetaSNP_TCICHK * MVresults_sort$BetaSNP_DM >= 0 ~ "same",
      MVresults_sort$BetaSNP_TCICHK * MVresults_sort$BetaSNP_DM < 0 ~ "different")

colPalette2 <- c("#00BFC4","gray", "#F8766D")

## Plot Manhattan with covariance colored
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color=effectdir), size=0.5) +
  scale_color_manual(values=colPalette2) +
  #scale_color_gradient2(low = "blue", mid="gray", high = "red", midpoint = 0, name="SNP effect covariance") +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes")) +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  geom_vline(aes(xintercept=0), lty="dashed", color="gray", linewidth=0.5) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP
#PSY2 v6 is 24153419 - 24156720

ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_revised_unrounded_weffectdir_", format(Sys.time(), "%m%d_%H%M"), ".jpg"), width=12, height=5)

ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_revised_unrounded_weffectdir_", format(Sys.time(), "%m%d_%H%M"), "_wide.jpg"), width=8, height=2)


table((MVresults_sort %>% filter(P < 0.001))$effectdir)

table((MVresults_sort %>% filter(P < Gao.cutoff))$effectdir)


## Plot Manhattan with SNP effect error covariance colored
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color=BetaSNP_TCICHKDM.cov), size=0.5) +
  #scale_color_manual(values=colPalette2) +
  scale_color_gradient2(low = "blue", mid="gray", high = "red", midpoint = 0, name="SNP effect covariance") +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes")) +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  geom_vline(aes(xintercept=0), lty="dashed", color="gray", linewidth=0.5) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP


## Plot Manhattan with residual (polygenic) genetic covariance colored
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color= resid_gen_cov), size=0.5) +
  scale_color_manual(values=colPalette2) +
  scale_color_gradient2(high = "blue", low="gray", mid = "lightblue", midpoint = -0.4, limits=c(-0.6,-0.19),
                        name="residual \ngenetic \ncovariance") +
 #scale_color_gradient(high = "blue", low="gray", name="residual polygenic covariance") +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes")) +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  #geom_vline(aes(xintercept=0), lty="dashed", color="gray", linewidth=0.5) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP
#PSY2 v6 is 24153419 - 24156720


#note: the residual genetic covariance gets smaller (less negative) when the SNP effects the traits in different directions (i.e. the SNP is explaining some of the covariance)
#hist((MVresults_sort %>% filter(effectdir=="same"))$TCDMgencov)
#hist((MVresults_sort %>% filter(effectdir=="different"))$TCDMgencov)
```

SNP contribution to negative genetic covariance:
```{r}
#BLUPs_hap_data <- readRDS(file=paste0(project_path, "data/BLUPs_hap_data_0826.RDS")) %>%
#  mutate(S1_24155522_dom = ifelse(S1_24155522 == 1, 1, 0))
#initG <- readRDS( file=paste0(project_path, "output/initG_traitgenvar_0826.RDS"))

#negative genetic covariance with no other SNPs or PSY2
    #model calculating SNP effect on each trait
   nullmodel <- asreml(data = BLUPs_PSY2_data,
        formula(paste0("cbind(TCICHK, DM) ~ trait  + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
        residual = ~units:corgh(trait), maxit=20)
   nullmodel_varcomp <- summary(nullmodel)$varcomp
    
nullmodel_TCICHKv <- nullmodel_varcomp["trait:vm(germplasmName, Ginverse)!trait_TCICHK", 1] + nullmodel_varcomp["units:trait!trait_TCICHK", 1]
nullmodel_DMv <- nullmodel_varcomp["trait:vm(germplasmName, Ginverse)!trait_DM", 1] + nullmodel_varcomp["units:trait!trait_DM", 1]
nullmodel_cov <- nullmodel_varcomp["trait:vm(germplasmName, Ginverse)!trait!DM:!trait!TCICHK.cor", 1] + nullmodel_varcomp["units:trait!trait!DM:!trait!TCICHK.cor", 1]
nullmodel_gcov <- nullmodel_varcomp["trait:vm(germplasmName, Ginverse)!trait!DM:!trait!TCICHK.cor", 1]


# null model with PSY2 
nullmodel_PSY2 <- asreml(data = BLUPs_PSY2_data,
        formula(paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
        residual = ~units:corgh(trait), maxit=20)
nullmodel_PSY2_varcomp <- summary(nullmodel_PSY2)$varcomp

nullmodel_PSY2_TCICHKv <- nullmodel_PSY2_varcomp["trait:vm(germplasmName, Ginverse)!trait_TCICHK", 1] + nullmodel_PSY2_varcomp["units:trait!trait_TCICHK", 1]
nullmodel_PSY2_DMv <- nullmodel_PSY2_varcomp["trait:vm(germplasmName, Ginverse)!trait_DM", 1] + nullmodel_PSY2_varcomp["units:trait!trait_DM", 1]
nullmodel_PSY2_cov <- nullmodel_PSY2_varcomp["trait:vm(germplasmName, Ginverse)!trait!DM:!trait!TCICHK.cor", 1] + nullmodel_PSY2_varcomp["units:trait!trait!DM:!trait!TCICHK.cor", 1]

#calculate AF of PSY2 allele:
# PSY2_imputed_rev <- readRDS(file="./data/PSY2KASP_revimputed_unrounded.RDS")
# PSY2_AF <- sum(PSY2_imputed_rev$S1_24155522)/(nrow(PSY2_imputed_rev)*2)
# 




#difference in polygenic covariance
# if the negative covariance goes up (more positive), then the SNP is contributing to the negative covariance
MVresults_sort$contributiontocov <- nullmodel_gcov - MVresults_sort$resid_gen_cov

## Plot Manhattan with residual (polygenic) genetic covariance colored
ggplot(MVresults_sort, aes(x=POS, y=-log10(P))) +
  geom_point(aes(color= contributiontocov), size=0.5) +
  #scale_color_manual(values=colPalette2) +
  scale_color_gradient2(high = "red", low="blue", mid = "gray", midpoint = 0.02,
                        name="contribution to \ngenetic covariance \nbetween TC and DM") +
 #scale_color_gradient(high = "blue", low="gray", name="residual polygenic covariance") +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM with PSY2 covariate, unrounded genotypes")) +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.4, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  #geom_vline(aes(xintercept=0), lty="dashed", color="gray", linewidth=0.5) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed") #PSY2 causal SNP
#PSY2 v6 is 24153419 - 24156720



ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_contributiontogencov_revised_unrounded_", format(Sys.time(), "%m%d_%H%M"), "_wide.jpg"), width=12, height=3)
```

Examine LD structure around top SNPs
```{r}
genomat_SNPs <- readRDS(file="./data/genomat_rev.RDS")

#starts with MVresults_sort from chunk above
all_SNP_results_df <- MVresults_sort

all_SNP_results_df$POS <- as.numeric(all_SNP_results_df$POS)
topSNPs <- all_SNP_results_df %>% filter(P < 0.001) #relaxed threshold

genomat_ROG <- genomat_t %>%
   mutate(CHROM= as.integer(gsub("S", "", SNP) %>% gsub("_[0-9].*", "", .)),
          POS= as.numeric(gsub("^S.*_", "", SNP)))


#flag peaks and calculate LD 
genomat_peaks_allchrom <- data.frame()
for(i in 1:18) {
  print(i)
  genomat_peaks_i <- genomat_ROG %>% filter(CHROM == i) 

  topSNPs_chri <- topSNPs %>% filter(CHROM == i) %>% arrange(P)
  peak1 <- topSNPs_chri[1,]

  #calculate LD between each SNP and the peak 1 SNP
  if(nrow(peak1) != 0 & !is.na(peak1$SNP_ID)){
    LDmatrix_peak1 <- genomat_peaks_i %>% dplyr::select(-c(SNP, compute_group, CHROM, POS)) %>% #drop metadata non-numeric columns
      t() %>% cor() #calculate correlation matrix
    LDwithpeak1 <- LDmatrix_peak1[,paste0(peak1$SNP_ID)] #pull out correlation with peak1 SNP
    genomat_peaks_i$LDpeak1 <- LDwithpeak1
    
    genomat_peaks_i <- genomat_peaks_i %>% mutate(distance_peak1 = POS - peak1$POS, peak1_SNP = peak1$SNP_ID)
  } else{genomat_peaks_i <- genomat_peaks_i %>% mutate(LDpeak1 = NA, distance_peak1 = NA, peak1_SNP = NA)}

  #check if there is another peak on this chromosome and repeat if so
  peak1_region <- c( peak1$POS - 2000000 , peak1$POS + 2000000)  
  genomat_around_peak1 <- genomat_peaks_i %>% 
       filter(POS > peak1_region[1] & POS < peak1_region[2])
  
  otherpeaks <- topSNPs_chri[!topSNPs_chri$SNP_ID %in% genomat_around_peak1$SNP,]
  peak2 <- otherpeaks %>% arrange(P) %>% head(n=1)
  
  #calculate LD between each SNP and the peak 2 SNP
  if(nrow(peak2) != 0){
    LDmatrix_peak2 <- genomat_peaks_i %>% dplyr::select(-c(SNP, compute_group, CHROM, POS, distance_peak1, peak1_SNP)) %>% 
      t() %>% cor()
    LDwithpeak2 <- LDmatrix_peak2[,paste0(peak2$SNP_ID)]
    genomat_peaks_i$LDpeak2 <- LDwithpeak2
    
    genomat_peaks_i <- genomat_peaks_i %>% mutate(distance_peak2 = POS - peak2$POS, peak2_SNP = peak2$SNP_ID)
  } else{genomat_peaks_i <- genomat_peaks_i %>% mutate(LDpeak2 = NA, distance_peak2 = NA, peak2_SNP = NA)
  } #end chromosome
  
  #join chromosome results with the others
 genomat_peaks_allchrom <- rbind(genomat_peaks_allchrom, genomat_peaks_i)
 
}

# Examine LD decay around top SNPs
LDdecay <- data.table::rbindlist(list(genomat_peaks_allchrom %>% dplyr::select(distance_peak1, LDpeak1),
                 genomat_peaks_allchrom %>% dplyr::select(distance_peak2, LDpeak2)), use.names=F) %>%
  filter(!is.na(distance_peak1)) %>%
  mutate(Rsquared = LDpeak1^2)

#Find average LD within 1 kb bins
bins=seq(1,max(LDdecay$distance_peak1, na.rm=T), by=1000)
my.means=rep(0, length(bins))
LD.averages=data.frame(bins, my.means)
for (i in 1:length(bins)) {
    data.interval=subset(LDdecay, (LDdecay$distance_peak1 >= bins[i] & LDdecay$distance_peak1 < (bins[i]+500)))
    LD.averages$my.means[i]=mean(data.interval$Rsquared)
}
plot(LD.averages$bins, LD.averages$my.means)
plot(LD.averages$bins, LD.averages$my.means, xlim=c(0, 80000))

# Visualize LD decay
ggplot(data= LDdecay, aes(x=abs(distance_peak1), y=Rsquared)) +
         geom_point() +
         geom_smooth(se = FALSE, method = "loess") +
         ggtitle("LD decay around peak SNPs") +
         xlab("distance (bp) to peak SNP") +
         ylab("correlation with peak SNP")

ggsave(filename=paste0(project_path, "figures/asreml_mvGWAS_LDdecay_unroundedgenos_",
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))


# Plot Manhattan colored by LD with peak SNP
all_SNP_results_peaks <- genomat_peaks_allchrom %>% dplyr::select(c(SNP, CHROM, POS, LDpeak1, distance_peak1, peak1_SNP, LDpeak2, distance_peak2, peak2_SNP)) %>% 
  mutate(CHROM = as.character(CHROM)) %>%
  mutate(distance_any = case_when(is.na(distance_peak2) | 
                                    abs(distance_peak1) < abs(distance_peak2) ~ distance_peak1, 
                                  TRUE ~ distance_peak2)) %>%
 mutate(Rsquared_any = case_when(is.na(distance_peak2) | 
                                   abs(distance_peak1) < abs(distance_peak2) ~ LDpeak1^2, 
                                 TRUE ~ LDpeak2^2)) %>%
  mutate(peak_SNP = case_when(SNP == peak1_SNP | SNP == peak2_SNP ~ "yes", T ~ "no")) %>%
  left_join(all_SNP_results_df, .) %>% 
  arrange(as.numeric(CHROM), as.numeric(POS))

#order factor levels so they display in order of chromosome
all_SNP_results_peaks$SNP <- factor(all_SNP_results_peaks$SNP, 
                                        levels = all_SNP_results_peaks$SNP)
all_SNP_results_peaks$CHROM <- as.factor(all_SNP_results_peaks$CHROM)
all_SNP_results_peaks$CHROM <- factor(all_SNP_results_peaks$CHROM, 
                                               levels = as.character(c(1:18)))

saveRDS(all_SNP_results_peaks, file=paste0(project_path, "output/all_SNP_results_MVWald_wPSY2KASPand5PC2_unrounded_annotations_effdir_", format(Sys.time(), "%m%d_%H%M"), ".RDS"))


## Plot Manhattan with joint p-values, colored by LD with peak SNP
ggplot(all_SNP_results_peaks, aes(x=POS, y=-log10(P), color=Rsquared_any)) +
  geom_point(aes(pch=peak_SNP, size=peak_SNP)) +
  scale_color_gradient2(low = "blue", mid="orange", high = "red", midpoint = 0.5, 
                          name= expression(textstyle("R"^2*" with starred SNP"))) +
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM, PSY2 covariate, unrounded\n with LD around peak SNPs")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.2) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.2, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
  geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed", lwd=0.2) #PSY2 causal SNP


ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_revised_unrounded_wLD_", format(Sys.time(), "%m%d_%H%M"), ".jpg"), width=12, height=5)


# Color Manhattan by effect direction same/different
colPalette2 <- c("#00BFC4","gray", "#F8766D")

ggplot(all_SNP_results_peaks, aes(x=POS, y=-log10(P), color=effectdir)) +
  geom_point(aes(pch=peak_SNP, size=peak_SNP)) +
  scale_color_manual(values = colPalette2, name="effect direction") +
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chromosome physical position") +
  ggtitle(paste0("MLMM GWAS for TC and DM, PSY2 covariate, unrounded\nwith effect direction")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.2) +
  geom_hline(yintercept = -log10(0.001), linetype="dashed", lwd=0.2, color="blue") +
  facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
    theme(strip.background = element_rect(fill = "white"),
          axis.text.x = element_blank(),  # Remove x-axis text
          axis.ticks.x = element_blank(),
          panel.spacing = unit(0, "lines")) +
   geom_vline(data = subset(MVresults_sort, CHROM=="1"), 
             aes(xintercept=24155522), color="orange", lty="dashed", lwd=0.2) #PSY2 causal SNP

ggsave(filename=paste0(project_path, "figures/Manhattan_mvGWASwPSY2KASPand5PCs_unrounded_wEffectDir_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))



# Binomial test
peakSNPsonly <- all_SNP_results_peaks %>% filter(peak_SNP == "yes")
nrow(peakSNPsonly) #15, with 10 out of 15 in opposite directions
binom.test(x=c(table(peakSNPsonly$effectdir)), p=0.5, alternative = "greater")
#p-value = 0.166

# Reattach allele Ref/Alt information and MAF:
genomat_SNPs <- readRDS(file="./data/genomat_rev.RDS")
genomat_SNPinfo <- readRDS(file=paste0(project_path, "data/AllChrom_MergedCASSYellowDARTGenotypesSharedSites_REF19imputedAndFiltered_SNPinfo.RDS")) %>% filter(ID %in% colnames(genomat_SNPs)) %>% rename("SNP_ID"="ID") 
genomat_SNPinfo$AF <- colSums(genomat_SNPs) / (nrow(genomat_SNPs)*2)


peakSNPsonly_alleles <- left_join(peakSNPsonly, genomat_SNPinfo %>% dplyr::select(SNP_ID,ALT,REF,AF)) %>%
  mutate(Minorallele = case_when(AF >= 0.5 ~ REF, AF < 0.5 ~ ALT),
         Majorallele = case_when(AF >= 0.5 ~ ALT, AF < 0.5 ~ REF)) %>%
  mutate(MAF = case_when(AF < 0.5 ~ AF, AF >= 0.5 ~ (1-AF) ))

table2 <- peakSNPsonly_alleles %>% arrange(P) %>% dplyr::select(SNP_ID,P,BetaSNP_TCICHK,BetaSNP_TCICHK.se, BetaSNP_DM,BetaSNP_DM.se,effectdir,REF,ALT,AF) 

# get total phenotypic variance for TCICHK and DM to calculate PVE
totalv_TCICHK <- 0.4393742 #var(modeldata$TCICHK, na.rm=T) 
totalv_DM <- 0.3920023 #var(modeldata$DM, na.rm=T) 

table2_rounded <- table2 %>%
  #calculate proportion variance explained (PVE)
  mutate(PVE_TCICHK = (BetaSNP_TCICHK^2 * 2 * AF * (1-AF))/ totalv_TCICHK,
         PVE_DM = (BetaSNP_DM^2 * 2 * AF * (1-AF))/ totalv_DM) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))

## TABLE 2
View(table2_rounded)

write.csv(table2_rounded, file=paste0(project_path, "output/table2_wPSY2KASPand5PCs_revised_unroundedgenotypes_", format(Sys.time(), "%m%d_%H%M"), ".csv"))


#total proportion covariance explained by the peak SNPs:
sum(table2_rounded$PCE_cov)


#total proportion covariance explained by the PSY2 SNPs:
sum(table2_rounded$PCE_cov)
```

Test significance of negative genetic covariance in model with PSY2 and all SNPs passing the relaxed threshold:
```{r}
PeakSNPsdata <- genomat_t %>%
  filter(genomat_t$SNP %in% peakSNPsonly$SNP_ID) 

rownames(PeakSNPsdata) <- PeakSNPsdata$SNP
PeakSNPsdata_t <- as.data.frame(t(PeakSNPsdata %>% dplyr::select(-c(SNP, compute_group))))
colnames(PeakSNPsdata_t) <- paste0(PeakSNPsdata$SNP)
PeakSNPsdata_t$germplasmName <- rownames(PeakSNPsdata_t)

modeldata <- left_join(BLUPs_PSY2_data, PeakSNPsdata_t, by="germplasmName") %>% 
      mutate(germplasmName = as.factor(germplasmName))
    
    
#fit model with all suggestively associated SNPs and PSY2 effect
model_allpeakSNPs <- asreml(data= modeldata, 
       fixed = formula (paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + ", 
                              paste0("PC", 1:5, collapse="+"), "+", 
                              paste0("trait:", peakSNPsonly$SNP_ID, collapse="+")) ),
       random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
       residual =  ~units:corgh(trait), maxit=20)

saveRDS(model_allpeakSNPs, file=paste0(project_path, "output/fullmodel_PSY2KASP_revised_unrounded_15peakSNPs.RDS"))

#fit reduced model with the residual genetic covariance set to 0

model_allpeakSNPs_0cov <- asreml(data = modeldata, 
       fixed= formula (paste0("cbind(TCICHK, DM) ~ trait + trait:S1_24155522 + ", 
                              paste0("PC", 1:5, collapse="+"), "+", 
                              paste0("trait:", peakSNPsonly$SNP_ID, collapse="+"))),
       random = ~ diag(trait):vm(germplasmName, Ginverse),
       residual =  ~units:diag(trait), maxit=20)

LRTresults <- lrt.asreml(model_allpeakSNPs_0cov, model_allpeakSNPs, boundary=T)
saveRDS(LRTresults, file=paste0(project_path, "output/LRT_residual_gen_covar_revised_unrounded.RDS"))



#LD between top SNPs:
peakSNPscormat <- cor( modeldata %>% dplyr::select(S1_24155522, peakSNPsonly$SNP_ID) )
peakSNPs_Rsquared <- peakSNPscormat^2
peakSNPs_Rsquared_PSY2 <- ((as.data.frame(peakSNPscormat)$S1_24155522)[-1])^2
max(peakSNPs_Rsquared_PSY2)
```
A significant likelihood ratio test here (p=0.005) indicates that the model with negative residual genetic covariance is a significantly better fit than one without it. We can conclude that there is significant negative residual genetic covariance even after accounting for all of the suggestively associated SNPs.


Test out adding weights to the model:
```{r}

nullmodel_TCICHKweights <- asreml(data = BLUPs_PSY2_data,
        formula(paste0("TCICHK ~ PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ vm(germplasmName, Ginverse),
        residual = ~units, maxit=20, weights=acc_weight)

nullmodel_biweights <- asreml(data = BLUPs_PSY2_data,
        formula(paste0("cbind(TCICHK,DM) ~ PC1 + PC2 + PC3 + PC4 + PC5")),
         random = ~ corgh(trait, init=initG):vm(germplasmName, Ginverse),
        residual = ~units:us(trait), maxit=20, weights=acc_weight)
```


#### Original BLUPs, re-imputed unrounded genotype data, WITH WEIGHTS
(Weights are mean of TCICHK and DM BLUP weights)
Combine results from the parallel computation sections: 
```{r}
project_path <- "/workdir/ssv42/mvGWAS/"
dataset <- "CASS"
ncores <- 60
lastrundate <- "0716" # may need a tryCatch if sections finished on different dates

all_SNP_weighted_results_df <- data.frame()
for (i in 1:ncores){
  section <- i
  print(i)
  
  lookupresults <- tryCatch(
    SNPresults_i <- readRDS(file=paste0(project_path,
               "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_weighted_section_", 
               section, "_", lastrundate, ".RDS")),
    error = function(e) e)

  # if(inherits(lookupresults, "error") ){
  #   print("trying previous day")
  #   lastrundate <- "1706"
  #    SNPresults_i <- readRDS(file=paste0(project_path,
  #              "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_section_", 
  #              section, "_", lastrundate, ".RDS"))
  # }

  all_SNP_weighted_results_df <- rbind(all_SNP_weighted_results_df, SNPresults_i)
}

# extract SNP information
all_SNP_weighted_results_df$SNP_ID <- rownames(all_SNP_weighted_results_df)
all_SNP_weighted_results_df$CHROM <- gsub("_[0-9].*", "",  rownames(all_SNP_weighted_results_df)) %>% gsub("S","",.)
all_SNP_weighted_results_df$POS <- gsub("S[0-9]*_", "",  rownames(all_SNP_weighted_results_df))


saveRDS(all_SNP_weighted_results_df, file=paste0(project_path,
             "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_weighted_allSNPresults_",
             format(Sys.time(), "%m%d"), ".RDS"))
```


