---
title: "07_GWAS_for_other_traits"
output: html_document
date: "2024-03-20"
---


# About
Run univariate trait associations using mixed linear models in asreml.

# Note:
Copy and paste the code below into the terminal version of R in a screen session (so that Rstudio remains free while the GWAS is running)

## Univariate GWAS
# Setup and load data
```{r}
library(tidyverse); library(asreml); library(aod)
project_path <- "/workdir/ssv42/mvGWAS/"
dataset <- "CASS"
ncores <- 60
  
### DEBUG settings -- test run one SNP model
DEBUG <- TRUE

#read in original drgBLUP phenotype data
BLUPs_data <- readRDS(file=paste0(project_path, "data/BLUPs_scaled_and_hap_data_", dataset, ".RDS")) 
  
# Read in revised genotype matrix 
genomat_SNPs <- readRDS(file="./data/genomat_rev.RDS")
#dim(genomat_SNPs) #377 x 43720
  
# Filter for common accessions
common_acc <- intersect(rownames(genomat_SNPs), BLUPs_data$germplasmName)
genomat_SNPs <- genomat_SNPs[common_acc,]
BLUPs_data <- BLUPs_data %>% filter(germplasmName %in% common_acc)
      
# Get (revised version) relationship matrix
Ginverse <- readRDS(file=paste0(project_path, "data/Ginverse_revisedimputation.RDS"))
Gsampleorder <- rownames(Ginverse) #make sure names are the same order
      
# Add computational group sections to SNPs
genomat_t <- t(genomat_SNPs) %>% as.data.frame()
genomat_t$SNP <- rownames(genomat_t)
nSNPs <- dim(genomat_t)[1]
genomat_t$compute_group <- as.numeric(cut_number(c(1:nSNPs), ncores))
      
# Read in imputed PSY2 KASP marker covariate
PSY2_imputed_rev <- readRDS(file=paste0(project_path, "data/PSY2KASP_revimputed_unrounded.RDS"))
      # Get original PSY2 KASP marker covariate
      # genomat_KASP_OG <- readRDS(file=paste0(project_path, "data/genomat_plusKASP_filtered.RDS"))
      #  PSY2marker_OG <- as.data.frame(t(genomat_KASP_OG[genomat_KASP_OG$SNP == "S1_24155522",] %>%
      #   dplyr::select(-c("SNP","ALT","REF","compute_group"))))
      #  colnames(PSY2marker_OG) <- "S1_24155522"
      #  PSY2marker_OG$germplasmName <- rownames(PSY2marker_OG)
      #  PSY2marker_OG$S1_24155522_dom <- ifelse(PSY2marker_OG$S1_24155522 == 1, 1, 0) 
    
#Get PC covariates
PCcovariates <- readRDS("/workdir/ssv42/mvGWAS/data/PCcovariates_rev.RDS") %>%
  as.data.frame() %>% rownames_to_column("germplasmName")
      # Get original PC covariates
       # PCaccessionnames <- read.table(file= paste0(project_path, "data/", dataset, "_accessionnames.txt"))[,1]
       # dosagesPC <- as.data.frame(readRDS(file=paste0(project_path,
       #"data/DosageMatrix_PCA_CASS_PCmatrix_copy.RDS")))[,1:5]
       # dosagesPC$germplasmName <- PCaccessionnames
      
#Get weights
BLUPs_revised <- readRDS(file=paste0(project_path,
                                     "data/alltrait_scaled_drgBLUPs_rev_CASS_TCICHK_H2filterFALSE.RDS")) %>%
  filter(germplasmName %in% Gsampleorder) %>%
  mutate(acc_weight = rowMeans(cbind(TCICHK_weight, DM_weight), na.rm = TRUE)) %>% 
  #use mean of the two traits for accession-level weights 
  mutate(across(contains("_weight"), ~ if_else(is.na(.x), min(.x, na.rm = TRUE), .x)))
  #fill any NAs with the lowest weight because there cannot be any missing values
BLUP_weights <- BLUPs_revised %>% dplyr::select(germplasmName, contains("_weight"))
      
#Merge BLUP and PSY2 marker data together, ensuring correct order of sample names
BLUPs_PSY2_data <- left_join(data.frame(germplasmName = Gsampleorder), BLUPs_data, by="germplasmName") %>%
  left_join(PSY2_imputed_rev, by="germplasmName") %>%
  left_join(PCcovariates, by="germplasmName") %>%
  left_join(BLUP_weights, by="germplasmName") %>%
  mutate(germplasmName = as.factor(germplasmName))

#confirm correct order
all(BLUPs_PSY2_data$germplasmName == Gsampleorder)
length(BLUPs_PSY2_data$germplasmName) == 376
nAccessions <- length(BLUPs_PSY2_data$germplasmName) 



### Define GWAS function

# Run univariate associations with TCICHK and DM with PSY2 SNP as fixed effect:
SNP_uniGWAS_function <- function(section, GWAStrait){
  
  library(tidyverse); library(asreml); library(aod)
  
  #select SNPs in this compute group
  section_SNPs <- genomat_t %>% filter(compute_group == section)
  
  #initiate storage for SNP effects
  all_SNP_results_df <- as.data.frame(matrix(nrow = nrow(section_SNPs), ncol = 7))
  colnames(all_SNP_results_df) <- c("Beta_SNP", "Beta_SNP_SE", "Zratio", "Df", "Sum_Sq", "Wald_stat", "Pr")
  
  SNPstotest <- seq_along(section_SNPs$SNP)
  if(DEBUG == TRUE){ SNPstotest <- 1 }
  
  for(i in SNPstotest){
    #debug note: chr 1 peak is at: i = 292; section = 1
    SNPi <- section_SNPs$SNP[i]
    SNPidata <- as.data.frame(t(section_SNPs[i, 1:nAccessions]))
    colnames(SNPidata) <- paste0(SNPi)
    SNPidata$germplasmName <- rownames(SNPidata)

    modeldata <- left_join(BLUPs_PSY2_data, SNPidata, by="germplasmName") %>% 
      mutate(germplasmName = as.factor(germplasmName)) %>%
      mutate(traiti_weights = .data[[paste0(GWAStrait, "_weight")]])
    
    #only add PSY2 as a covariate for the relevant TC & DM traits
    if(GWAStrait %in% c("TCICHK","TCHART","DM")){
      fixedformula <- formula(paste0(GWAStrait, " ~ 1 + ", SNPi, " + S1_24155522 + PC1 + PC2 + PC3 + PC4 + PC5"))
    } else {
      fixedformula <- formula(paste0(GWAStrait, " ~ 1 + ", SNPi, " + PC1 + PC2 + PC3 + PC4 + PC5"))
    }
    
    #univariate model with PCs
    SNPimodelwPCs <- asreml(data = modeldata,
       fixed = fixedformula,
         random = ~ vm(germplasmName, Ginverse),
        residual = ~ units, weights= traiti_weights)
    
    SNPeff_wPCs <- summary(SNPimodelwPCs, coef=T)$coef.fixed[paste0(SNPi),]
    
    #save model with PCs
    # SNPimodelwPCs <- SNPimodelwPCs
    # model_storage_PCs[[i]] <- SNPimodelwPCs
    # model_storage_PCs <<- model_storage_PCs
    
    #Wald test for significance of fixed effects
    #WaldTest <- wald.asreml(SNPimodelnoPCs)
    WaldTest_wPCs <- wald.asreml(SNPimodelwPCs)
    

    #bind together and save results to dataframe
    SNP_results_i <- cbind(t(SNPeff_wPCs), t(WaldTest_wPCs[paste0(SNPi),]))
    
    all_SNP_results_df[i,] <- SNP_results_i
    rownames(all_SNP_results_df)[i] <- SNPi #save SNP_ID
    
    saveRDS(all_SNP_results_df, file=paste0(project_path,
                  "output/univariate_models/", GWAStrait, "_all_SNP_results_wPCs_revised_weighted_section",
                  section, "_", format(Sys.time(), "%m%d"), ".RDS"))
    
    #save model object
    # saveRDS(model_storage_PCs, file=paste0(project_path,
    #               "output/univariate_models/", GWAStrait, "_modelstorage_wPCs_section",
    #               section, "_", dataset, "_",
    #           format(Sys.time(), "%m%d"), ".RDS"))
    
    cat(paste0(SNPi, ", ", i, " of ", length(section_SNPs$SNP), " in section ", section, "\n"),
      file = paste0(project_path, "output/univariate_models/progressbar/", GWAStrait, "_PSY2KASPand5PCs_weighted_run_", format(Sys.time(), "%m%d"), ".txt"), append=T)

  } #end SNP

} #end compute section
```

Run GWAS function in parallel
```{r, eval=F}
#Run asreml GWAS function over multiple cores:
library(parallel)
DEBUG <- FALSE
sections <- 1:ncores

# name functions and data to export to cluster
GWASfunctionvariables <- c("SNP_uniGWAS_function", "genomat_t", "Ginverse", "BLUPs_PSY2_data", 
                           "project_path", "ncores", "DEBUG", "nAccessions")
  
# create cluster object
cl <- makeCluster(ncores)
clusterExport(cl, varlist = GWASfunctionvariables)

# test each SNP in chromosome 
results1 <- parSapply(cl, sections, GWAStrait = "TCICHK", SNP_uniGWAS_function)
results2 <- parSapply(cl, sections, GWAStrait = "TCHART", SNP_uniGWAS_function)
results3 <- parSapply(cl, sections, GWAStrait = "DM", SNP_uniGWAS_function)
results5 <- parSapply(cl, sections, GWAStrait = "RTWT", SNP_uniGWAS_function)
results6 <- parSapply(cl, sections, GWAStrait = "HI", SNP_uniGWAS_function)
results7 <- parSapply(cl, sections, GWAStrait = "SPRTPERC", SNP_uniGWAS_function)
results8 <- parSapply(cl, sections, GWAStrait = "FRYLD", SNP_uniGWAS_function)
results9 <- parSapply(cl, sections, GWAStrait = "CMD3S", SNP_uniGWAS_function)
results10 <- parSapply(cl, sections, GWAStrait = "CMD3I", SNP_uniGWAS_function)
results4 <- parSapply(cl, sections, GWAStrait = "DYLD", SNP_uniGWAS_function)

# close cluster object
stopCluster(cl)
```

Combine results from parallel computation sections for each trait one at a time:
```{r}
library(tidyverse)
project_path <- "/workdir/ssv42/mvGWAS/"
ncores <- 60
lastrundate <- "0717" # may need a tryCatch if sections finished on different dates
PCstatus <- "wPCs" #noPCs or wPCs

GWAStraitlist <- c("TCICHK","TCHART","DM","DYLD","RTWT","HI", "SPRTPERC","FRYLD","CMD3S","CMD3I")

for(j in seq_along(GWAStraitlist)){
  
GWAStrait <- GWAStraitlist[j]
print(GWAStrait)

    all_SNP_results_df <- data.frame()
    for (i in 1:ncores){
      section <- i
      #print(i)
      
      lookupresults <- tryCatch(
        SNPresults_i <- readRDS(file=paste0(project_path,
                   "output/univariate_models/", GWAStrait, "_all_SNP_results_", PCstatus,
                   "_revised_weighted_section", section, "_", lastrundate, ".RDS")),
        error = function(e) e)
    
      all_SNP_results_df <- rbind(all_SNP_results_df, SNPresults_i)
    }
    
    # extract SNP information
    all_SNP_results_df$SNP_ID <- rownames(all_SNP_results_df)
    all_SNP_results_df$CHROM <- gsub("_[0-9].*", "",  rownames(all_SNP_results_df)) %>% gsub("S","",.)
    all_SNP_results_df$POS <- gsub("S[0-9]*_", "",  rownames(all_SNP_results_df))
    
    saveRDS(all_SNP_results_df, file=paste0(project_path,
                 "output/univariate_models/", GWAStrait, "_all_SNP_results_df_revised_weighted_", 
                 format(Sys.time(), "%m%d"), ".RDS"))
} #end trait

```


## Examine the results
Make Manhattan and Q-Q plots
```{r}
lastrundate <- "0717"
GWAStraitlist <- c("TCICHK","TCHART","DM","DYLD","RTWT","HI", "SPRTPERC","FRYLD","CMD3S","CMD3I")

for(j in seq_along(GWAStraitlist)){
  
  GWAStrait <- GWAStraitlist[j]

  traiti_SNP_results_df <- readRDS(file=paste0(project_path,
             "output/univariate_models/", GWAStrait, "_all_SNP_results_df_revised_weighted_", 
             lastrundate, ".RDS"))
  
  #print QQ plot in console
  #qqman::qq(traiti_SNP_results_df$Pr, main=paste0(GWAStrait))
  
  #save QQ plot image
  png(paste0(project_path, "figures/", GWAStrait, "_revised_QQplot_wPCs_",
                      format(Sys.time(), "%m%d"), ".png"), 
          width=6, height=6, res=250, units="in")
   
    qqman::qq(traiti_SNP_results_df$Pr, main=paste0(GWAStrait))
  
  dev.off()


  #Make Manhattan plots:
  
  # Set significance cut-off: Gao-modified Bonferroni
  Gao.cutoff <- 1.367e-05
  
  traiti_SNP_results_df$POS <- as.numeric(traiti_SNP_results_df$POS)
  traiti_results <- traiti_SNP_results_df %>% arrange(as.numeric(CHROM), as.numeric(POS))
  
  #order factor levels so they display in order of chromosome
  traiti_results$SNP_ID <- factor(traiti_results$SNP_ID, levels = traiti_results$SNP_ID)
  traiti_results$CHROM <- as.factor(traiti_results$CHROM)
  traiti_results$CHROM <- factor(traiti_results$CHROM, levels = as.character(c(1:18)))
  
  #set color scheme
  colPalette <- rep(c("blue4","pink3"),9)
  
  ## Plot Manhattan with bivariate Wald test p-values
  ggplot(traiti_results, aes(x=POS, y=-log10(Pr))) +
    geom_point(aes(color=CHROM), size=0.5) +
    scale_color_manual(values=colPalette) +
    labs(y="-log10(p)", x="Chromosome physical position") +
    ggtitle(paste0("Univariate GWAS for ", GWAStrait, " with 5 PCs, weighted")) +
    guides(color="none")+
    geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
    facet_grid(. ~ CHROM, scales = "free_x", switch="x") +
      theme(legend.position = "none",
            strip.background = element_rect(fill = "white"),
            axis.text.x = element_blank(),  # Remove x-axis text
            axis.ticks.x = element_blank(),
            panel.spacing = unit(0, "lines")) +
    geom_vline(data = subset(traiti_results, CHROM=="1"), 
               aes(xintercept=24155522), color="orange", lty="dashed", 
               linewidth=if(GWAStrait %in% c("TCICHK","TCHART","DM")) 0.25 else 0) 
  #add PSY2 causal SNP annotation only if it was used as a covariate in the model
  
  
  ggsave(filename=paste0(project_path, "figures/", GWAStrait, "_GWAS_revised_weighted_",
                         format(Sys.time(), "%m%d_%H%M"), ".jpg"), width=12, height=5)

  if( GWAStrait == "TCICHK"){
    saveRDS(traiti_results, file=paste0(project_path, "output/TCICHK_univariate_results_revised_weighted.RDS")) }
  if( GWAStrait == "DM"){
    saveRDS(traiti_results, file=paste0(project_path, "output/DM_univariate_results_revised_weighted.RDS")) }

}
```


Zoom in on chromosome 1 QTL region for TCICHK and DM:
Chr. 1 TCICHK
```{r}
TCICHK_results <- readRDS(file=paste0(project_path, "output/TCICHK_univariate_results_revised_weighted.RDS"))
chr1_results <- TCICHK_results %>% filter(CHROM == 1)

#genome matrix (after MAF filtering)
genomat <- readRDS(file=paste0(project_path, "data/genomat_t_revised.RDS")) %>%
  mutate(CHROM =  gsub("_[0-9].*", "",  SNP) %>% gsub("S","",.),
         POS = as.numeric(sub("S[0-9]*_", "", SNP)))

genomat_QTLregion <- genomat %>% filter(CHROM=="1") # & POS > 20000000 & POS < 30000000
rownames(genomat_QTLregion) <- genomat_QTLregion$SNP

# Calculate LD with peak SNPs:
#flag peaks and calculate LD 
genomat_peaks_LD <- data.frame(SNP = genomat_QTLregion$SNP)

topSNPs_chr1 <- chr1_results %>% filter(Pr < relaxed.cutoff) %>% arrange(Pr)

#calculate LD between each SNP and the peak 1 SNP
peak1 <- topSNPs_chr1[1,]
  LDmatrix_chr1 <- genomat_QTLregion %>% dplyr::select(-c(SNP, compute_group, CHROM, POS)) %>% #drop metadata non-numeric columns
      t() %>% cor() #calculate correlation matrix
    LDwithpeak1 <- LDmatrix_chr1[,paste0(peak1$SNP_ID)] #pull out correlation with peak1 SNP
    genomat_peaks_LD$LDpeak1 <- LDwithpeak1
    genomat_peaks_LD$Rsquared_peak1 <- genomat_peaks_LD$LDpeak1^2
    
#calculate LD between each SNP and the peak 2 SNP
peak2 <- topSNPs_chr1[2,]
    LDwithpeak2 <- LDmatrix_chr1[,peak2$SNP_ID]
    genomat_peaks_LD$LDpeak2 <- LDwithpeak2
    genomat_peaks_LD$Rsquared_peak2 <- genomat_peaks_LD$LDpeak2^2

genomat_QTL_LD <- right_join(genomat_peaks_LD, genomat_QTLregion, by="SNP") %>%
  mutate(peak1SNP = as.numeric(gsub("S1_", "", peak1$SNP)), peak2SNP = as.numeric(gsub("S1_", "", peak2$SNP))) %>%
  mutate(distance_peak1 = POS - peak1SNP, distance_peak2 = POS - peak2SNP) %>%
  mutate(isthispeak1 = case_when(SNP == peak1$SNP_ID ~ "yes", T ~ "no")) %>%
  mutate(isthispeak2 = case_when(SNP == peak2$SNP_ID ~ "yes", T ~ "no")) %>%
  dplyr::select(SNP, POS, CHROM, peak1SNP, peak2SNP, distance_peak1, distance_peak2, Rsquared_peak1, Rsquared_peak2, isthispeak1, isthispeak2) %>%
  right_join(chr1_results %>% mutate(SNP=as.character(SNP_ID))) %>%
  arrange(POS)

## Plot chr 1 Manhattan with joint p-values, colored by LD with peak 1 SNP
ggplot(genomat_QTL_LD %>% filter(POS > 20000000), aes(x=POS, y=-log10(Pr), color=Rsquared_peak1)) +
  geom_point(aes(pch=isthispeak1, size=isthispeak1)) +
  scale_color_gradient2(low = "blue", mid="orange", high = "red", midpoint = 0.5, name=
                          expression(textstyle("R"^2*" with starred SNP")))+
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chr. 1 physical position (bp)") +
  ggtitle(paste0("TCICHK GWAS chr. 1 with PSY2, 5 PCs, weighted")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  theme(strip.background = element_rect(fill = "white")) +
   annotate("segment", x = c(19400000, 19600000), xend = c(19600000, 19800000), y = c(-0.1, -0.1), yend = c(0.1, 0.1)) +
  coord_cartesian(clip = "off", xlim = c(19000000, 35000000)) +
  geom_vline(aes(xintercept=24155522), color="orange", lty="dashed") #causal SNP position

ggsave(filename=paste0(project_path, "figures/TCICHK_GWAS_Chr1_wPSY2KASPand5PCs_revised_weighted_Manhattan_wLD_peak1_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))

ggsave(filename=paste0(project_path, "figures/TCICHK_GWAS_Chr1_wPSY2KASPand5PCs_revised_weighted_Manhattan_wLD_peak1_wide_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"), height=5, width=8, units="in")


ggplot(genomat_QTL_LD %>% filter(POS > 20000000), aes(x=POS, y=-log10(Pr), color=Rsquared_peak2)) +
  geom_point(aes(pch=isthispeak2, size=isthispeak2)) +
  scale_color_gradient2(low = "blue", mid="orange", high = "red", midpoint = 0.5, name=
                          expression(textstyle("R"^2*" with starred SNP")))+
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chr. 1 physical position (bp)") +
  ggtitle(paste0("TCICHK GWAS chr. 1 with PSY2 KASP, 5 PCs, weighted")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  theme(strip.background = element_rect(fill = "white")) +
   annotate("segment", x = c(19400000, 19600000), xend = c(19600000, 19800000), y = c(-0.1, -0.1), yend = c(0.1, 0.1)) +
  coord_cartesian(clip = "off", xlim = c(19000000, 35000000)) +
  geom_vline(aes(xintercept=24155522), color="orange", lty="dashed") #causal SNP position

ggsave(filename=paste0(project_path, "figures/TCICHK_GWAS_Chr1_wPSY2KASPand5PCs_revised_weighted_Manhattan_wLD_peak2_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))



### Calculate LD with PSY2: ###
genomat_plusPSY <- readRDS(file="./data/genomat_rev_plusPSY2KASP.RDS") %>% t() %>% as.data.frame() %>%
  rownames_to_column("SNP") %>%
  mutate(CHROM =  gsub("_[0-9].*", "",  SNP) %>% gsub("S","",.),
         POS = as.numeric(sub("S[0-9]*_", "", SNP)))

LDPSY2_chr1 <- genomat_plusPSY %>% filter(CHROM=="1") %>% dplyr::select(-c(SNP, CHROM, POS)) %>% #drop metadata non-numeric columns
      t() %>% cor() #calculate correlation matrix


topSNPs_chr1 <- chr1_results %>% filter(Pr < relaxed.cutoff) %>% arrange(Pr)

#calculate LD between each SNP and the peak 1 SNP
peak1 <- topSNPs_chr1[1,]
  LDmatrix_chr1 <- genomat_QTLregion %>% dplyr::select(-c(SNP, compute_group, CHROM, POS)) %>% #drop metadata non-numeric columns
      t() %>% cor() #calculate correlation matrix
    LDwithpeak1 <- LDmatrix_chr1[,paste0(peak1$SNP_ID)] #pull out correlation with peak1 SNP
    genomat_peaks_LD$LDpeak1 <- LDwithpeak1
    genomat_peaks_LD$Rsquared_peak1 <- genomat_peaks_LD$LDpeak1^2
    
```

Chr. 1 DM 
```{r}
DM_results <- readRDS(file=paste0(project_path, "output/DM_univariate_results_revised_weighted.RDS"))

chr1_results <- DM_results %>% filter(CHROM == 1)

#flag peaks and calculate LD 
genomat_peaks_LD <- data.frame(SNP = genomat_QTLregion$SNP)

topSNPs_chr1 <- chr1_results %>% filter(Pr < 0.001) %>% arrange(Pr)

#calculate LD between each SNP and the peak 1 SNP
peak1 <- topSNPs_chr1[1,]
  # LDmatrix_chr1 <- genomat_QTLregion %>% dplyr::select(-c(SNP, POS, CHROM, ALT, REF)) %>% #drop metadata non-numeric columns
  #     t() %>% cor() #calculate correlation matrix
    LDwithpeak1 <- LDmatrix_chr1[,paste0(peak1$SNP_ID)] #pull out correlation with peak1 SNP
    genomat_peaks_LD$LDpeak1 <- LDwithpeak1
    genomat_peaks_LD$Rsquared_peak1 <- genomat_peaks_LD$LDpeak1^2
    
#calculate LD between each SNP and the peak 2 SNP
peak2 <- topSNPs_chr1[2,]
    LDwithpeak2 <- LDmatrix_chr1[,peak2$SNP_ID]
    genomat_peaks_LD$LDpeak2 <- LDwithpeak2
    genomat_peaks_LD$Rsquared_peak2 <- genomat_peaks_LD$LDpeak2^2

genomat_QTL_LD <- right_join(genomat_peaks_LD, genomat_QTLregion, by="SNP") %>%
  mutate(peak1SNP = as.numeric(gsub("S1_", "", peak1$SNP)), peak2SNP = as.numeric(gsub("S1_", "", peak2$SNP))) %>%
  mutate(distance_peak1 = POS - peak1SNP, distance_peak2 = POS - peak2SNP) %>%
  mutate(isthispeak1 = case_when(SNP == peak1$SNP_ID ~ "yes", T ~ "no")) %>%
  mutate(isthispeak2 = case_when(SNP == peak2$SNP_ID ~ "yes", T ~ "no")) %>%
  dplyr::select(SNP, POS, CHROM, peak1SNP, peak2SNP, distance_peak1, distance_peak2, Rsquared_peak1, Rsquared_peak2, isthispeak1, isthispeak2) %>%
  right_join(chr1_results %>% mutate(SNP=as.character(SNP_ID))) %>%
  arrange(POS)


## Plot chr 1 Manhattan with joint p-values, colored by LD with peak 1 SNP
ggplot(genomat_QTL_LD %>% filter(POS > 20000000), aes(x=POS, y=-log10(Pr), color=Rsquared_peak1)) +
  geom_point(aes(pch=isthispeak1, size=isthispeak1)) +
  scale_color_gradient2(low = "blue", mid="orange", high = "red", midpoint = 0.5, name=
                          expression(textstyle("R"^2*" with starred SNP"))) +
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chr. 1 physical position (bp)") +
  ggtitle(paste0("DM GWAS chr. 1 with PSY2, 5 PCs, weighted")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  theme(strip.background = element_rect(fill = "white")) +
   annotate("segment", x = c(19400000, 19600000), xend = c(19600000, 19800000), y = c(-0.1, -0.1), yend = c(0.1, 0.1)) +
  coord_cartesian(clip = "off", xlim = c(19000000, 35000000)) +
  geom_vline(aes(xintercept=24155522), color="orange", lty="dashed") #causal SNP position

ggsave(filename=paste0(project_path, "figures/DM_GWAS_Chr1_wPSY2KASPand5PCs_revised_weighted_Manhattan_wLD_peak1_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))

ggsave(filename=paste0(project_path, "figures/DM_GWAS_Chr1_wPSY2KASPand5PCs_revised_weighted_Manhattan_wLD_peak1_wide_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"), height=5, width=8, units="in")


## Plot chr 1 Manhattan with joint p-values, colored by LD with peak 1 SNP
ggplot(genomat_QTL_LD %>% filter(POS > 20000000), aes(x=POS, y=-log10(Pr), color=Rsquared_peak2)) +
  geom_point(aes(pch=isthispeak2, size=isthispeak2)) +
  scale_color_gradient2(low = "blue", mid="orange", high = "red", midpoint = 0.5, name=
                          expression(textstyle("R"^2*" with starred SNP"))) +
  scale_shape_manual(values = c(19,8)) +
  scale_size_manual(values=c(0.8,1.2)) +
  labs(y="-log10(p)", x="Chr. 1 physical position (bp)") +
  ggtitle(paste0("DM GWAS chr. 1 with PSY2, 5 PCs, weighted")) +
  guides(pch="none", size="none") +
  geom_hline(yintercept = -log10(Gao.cutoff), linetype="dashed", lwd=0.4) +
  theme(strip.background = element_rect(fill = "white")) +
   annotate("segment", x = c(19400000, 19600000), xend = c(19600000, 19800000), y = c(-0.1, -0.1), yend = c(0.1, 0.1)) +
  coord_cartesian(clip = "off", xlim = c(19000000, 35000000)) +
  geom_vline(aes(xintercept=24155522), color="orange", lty="dashed") #causal SNP position

ggsave(filename=paste0(project_path, "figures/DM_GWAS_Chr1_wPSY2KASPand5PCs_revised_weighted_Manhattan_wLD_peak2_", 
                       format(Sys.time(), "%m%d_%H%M"), ".jpg"))
```

