---
title: "06_candidate_gene_analysis"
output: html_document
date: "2024-03-13"
---

# About
TBD

# Setup
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse); library(qqman); library(asreml); library(parallel)
project_path <- "/workdir/ssv42/mvGWAS/"
setwd(project_path)
set.seed(14850)

#Select dataset
dataset <- "CASS" #"HP" #"CASS"
```

## Import list of all cassava genes from Phytozome and a priori candidate list
# Map from v4.1 transcripts to v6.1 gene names
# Transcript and gene names came from: Wang et al. 2014, Sreekumar et al. 2022, and Saithong et al. 2013
```{r}
transcript_gene_key <- read_delim(paste0(project_path, "data/Phytozome_gene_annotations/Mesculenta_305_v6.1.locus_transcript_name_map.txt"), delim="\t", skip=1, show_col_types = FALSE)
#warning is okay: it appropriately fills in NAs where necessary

gene_annotations <- read_delim(paste0(project_path, "data/Phytozome_gene_annotations/Mesculenta_305_v6.1.annotation_info.txt"), show_col_types = FALSE)

gff_genes <- ape::read.gff(paste0(project_path, "data/Phytozome_gene_annotations/Mesculenta_305_v6.1.gene.gff3")) %>%
  filter(type=="gene") %>%
  mutate(geneID = str_match(attributes, "Name=([^;]+)")[,2]) %>%
  dplyr::select(-c("source","score","phase"))

allgenes <- read_delim(paste0(project_path, "data/Mesculenta_v6_genes.tsv"), skip = 1, delim="\t",
                       col_names=c("gene_ID", "gene_synonym", "duplicate", "transcript_ID", "function_annot", "species", "chromosome", "start_pos", "end_pos"), show_col_types = FALSE) %>%
  dplyr::select(-c(duplicate, species)) %>%
  full_join(transcript_gene_key, by=c("gene_ID"="#new-locusName", "gene_synonym"="old-locusName", "transcript_ID"="new-transcriptName")) %>%
  full_join(gff_genes, by=c("gene_ID"="geneID")) %>%
  full_join(gene_annotations, by=c("gene_ID"="locusName", "transcript_ID"="transcriptName")) %>%
  mutate(start_pos = case_when(is.na(start_pos) ~ start, T ~ start_pos),
         end_pos = case_when(is.na(end_pos) ~ end, T ~ end_pos)) %>%
  mutate(chrom = as.integer(gsub("Chromosome0", "", chromosome) %>% gsub("Chromosome", "", .)))



apriori <- read.csv(paste0(project_path, "data/APrioriCandidateGenesTable.csv")) %>%
  dplyr::select(-c(Citations, Cellular.component, Genes.with..related.TFBS)) %>%
  mutate(transcript_to_gene = paste0(transcript.ID..Phytozome., ".g"))

matchbytranscript  <- apriori %>%
  left_join(allgenes, by=c("transcript.ID..Phytozome." = "old-transcriptName"), relationship = "many-to-many") 

matchbytranscriptfuzzy  <- apriori %>%
  left_join(allgenes, by=c("transcript_to_gene" = "gene_synonym"), relationship = "many-to-many") 

matchbygeneID  <- apriori %>%
  left_join(allgenes, by=c("Cassava.gene.name.s." = "gene_ID"), relationship = "many-to-many") %>%
  mutate(gene_ID = Cassava.gene.name.s.)

apriori_genes <- full_join(matchbytranscript, matchbygeneID) %>%
  full_join(matchbytranscriptfuzzy) %>%
  filter(!duplicated(gene_ID)) %>%
  mutate(chrom = as.integer(gsub("Chromosome", "", chromosome))) %>%
  filter(!is.na(chromosome))
```
(Warnings here about NA generation & parsing failures is OK.)


# Read in SNP hit data from bivariate analyses:
```{r}
all_SNP_results_df <- readRDS(file=paste0(project_path,
            "output/mvGWAS_asreml_output/mvGWASwPSY2KASPand5PCs_unrounded_weighted_allSNPresults_0716.RDS"))


all_SNP_results_peaks <- readRDS(file=paste0(project_path, "output/all_SNP_results_MVWald_wPSY2KASPand5PC2_unrounded_weighted_annotations_effdir_0717_1703.RDS"))
```


# Check if any a priori candidate genes were hit
```{r}
# all_SNP_results_peaks <- readRDS(file=paste0(project_path, "output/all_SNP_results_MVWald_noPCs_wpeak_annotations_effdir_0614_1512.RDS")) %>% 
#   filter(peak1_SNP %in% c("S1_24309741", "S10_2262535") | peak2_SNP == "S1_26700297") 
Bonferonni_passed_SNPs <- all_SNP_results_peaks %>% filter(peak_SNP == "yes" & P < Gao.cutoff)


all_SNP_results_peaks_highLD <- all_SNP_results_peaks %>%
  filter(peak1_SNP %in% Bonferonni_passed_SNPs$SNP) %>%
  filter(Rsquared_any > 0.75) 
  #filter(peak1_SNP %in% c("S1_26700297", "S10_2262535")) %>% #selecting just bonferroni threshold passed SNPs

dist_threshold <- 100000
threshold_chrom1 <- 375000
threshold_chrom4 <- 500000

#search whether a SNP in high LD with the peak SNP is within the distance threshold of the a priori gene candidate
apriorigene_hits <- data.frame()
for(i in 1:nrow(apriori_genes)){
  
  if(apriori_genes$chromosome[i] == "Chromosome01"){dist_threshold <- threshold_chrom1}
  if(apriori_genes$chromosome[i] == "Chromosome04"){dist_threshold <- threshold_chrom4}
  
  start_search <- apriori_genes$start_pos[i] - dist_threshold
  end_search <- apriori_genes$end_pos[i] + dist_threshold
  chrom_i <- apriori_genes$chrom[i]
  
  lookup <- all_SNP_results_peaks_highLD %>% 
    filter(CHROM == chrom_i & POS > start_search & POS < end_search) %>%
    arrange(-LDpeak1)
  
  if(nrow(lookup) != 0) {
    print(i)
    this_hit <- c(apriori_genes$Description[i], apriori_genes$chromosome[i],
                  apriori_genes$start_pos[i], apriori_genes$end_pos[i], 
                  lookup$LDpeak1[1], as.character(lookup$peak1_SNP[1]), lookup$P[1], lookup$distance_peak1[1])
    
    apriorigene_hits <- rbind(apriorigene_hits, this_hit)
    }
}
colnames(apriorigene_hits) <- c("gene", "chromosome", "start_pos", "end_pos", "LDwpeakSNP", "peakSNP", "peakSNPp", "distance_peaksnp")


print(apriorigene_hits)
#none are within 100kb cutoff 
```

#Old method: Get set of candidate genes near top SNPs using SNP-specific distance threshold:
```{r, eval=F}
candidate_genes <- data.frame()
for(i in 1:nrow(Bonferonni_passed_SNPs)){ #search genes
  
  chrom_i <- Bonferonni_passed_SNPs$CHROM[i]
  pos_i <- Bonferonni_passed_SNPs$POS[i]
  snp_i <- Bonferonni_passed_SNPs$SNP[i]
  
  #Find LD decay interval for this SNP
  genomat_around_peaki <- genomat_ROG %>% filter(CHROM == chrom_i & POS > pos_i - 500000 & POS < pos_i + 500000) %>%
    mutate(distance_peaki = abs(POS - pos_i))
  rownames(genomat_around_peaki) <- genomat_around_peaki$SNP

  LDmatrix_around_peaki <- genomat_around_peaki %>% dplyr::select(-c(SNP, compute_group, CHROM, POS, distance_peaki)) %>% 
        t() %>% cor()
  
      R2withpeaki <- (LDmatrix_around_peaki[,paste0(snp_i)])^2
      genomat_around_peaki$R2peaki <- R2withpeaki
      
      
      ggplot(genomat_around_peaki, aes(x=distance_peaki, y=R2peaki)) + geom_point() + xlim(c(0,100000))
      
  
  
  
  allgenes_chrom_i <- allgenes %>% filter(chrom == chrom_i)
  for(j in 1:nrow(allgenes_chrom_i)){
    
    start_search <- allgenes_chrom_i$start_pos[j] - dist_threshold
    end_search <- allgenes_chrom_i$end_pos[j] + dist_threshold
    
    if( pos_i >= start_search && pos_i <= end_search){
      
      gene_midpoint = (as.numeric(allgenes_chrom_i$start_pos[j]) + as.numeric(allgenes_chrom_i$end_pos[j]))/2
    
      this_candidate <- c(as.character(Bonferonni_passed_SNPs$SNP[i]),
                          allgenes_chrom_i$gene_ID[j],
                          allgenes_chrom_i$function_annot[j],
                          allgenes_chrom_i$GO[j],
                          allgenes_chrom_i$chrom[j],
                          allgenes_chrom_i$start_pos[j],
                          allgenes_chrom_i$end_pos[j],
                          abs(as.numeric(gsub("S1.*_", "", 
                                              as.character(Bonferonni_passed_SNPs$SNP[i]) )) - gene_midpoint))
      
      candidate_genes <- rbind(candidate_genes, this_candidate)
      
    }
  } 
}
colnames(candidate_genes) <- c("SNP", "gene_ID", "functional_annotation", "GOterm", "chrom", "startpos", "endpos", "mid_distance")
View(unique(candidate_genes))
```



#Newer method: using chromosome-specific distance threshold:
```{r}
dist_threshold <- 100000
threshold_chrom1 <- 100000 #375000
threshold_chrom4 <- 100000 #500000

candidate_genes <- data.frame()
for(i in 1:nrow(Bonferonni_passed_SNPs)){ #search genes
  
  chrom_i <- Bonferonni_passed_SNPs$CHROM[i]
  pos_i <- Bonferonni_passed_SNPs$POS[i]
  
  if(chrom_i == 1){
    dist_threshold <- threshold_chrom1} 
  if(chrom_i == 4){ 
    dist_threshold <- threshold_chrom4}
  
  allgenes_chrom_i <- allgenes %>% filter(chrom == chrom_i)
  for(j in 1:nrow(allgenes_chrom_i)){
    
    start_search <- allgenes_chrom_i$start_pos[j] - dist_threshold
    end_search <- allgenes_chrom_i$end_pos[j] + dist_threshold
    
    if( pos_i >= start_search && pos_i <= end_search){
      
      gene_midpoint = (as.numeric(allgenes_chrom_i$start_pos[j]) + as.numeric(allgenes_chrom_i$end_pos[j]))/2
    
      this_candidate <- c(as.character(Bonferonni_passed_SNPs$SNP[i]),
                          allgenes_chrom_i$gene_ID[j],
                          allgenes_chrom_i$function_annot[j],
                          allgenes_chrom_i$`arabi-defline`[j],
                          allgenes_chrom_i$GO[j],
                          allgenes_chrom_i$chrom[j],
                          allgenes_chrom_i$start_pos[j],
                          allgenes_chrom_i$end_pos[j],
                          abs(as.numeric(gsub("S1.*_", "", 
                                              as.character(Bonferonni_passed_SNPs$SNP[i]) )) - gene_midpoint))
      
      candidate_genes <- rbind(candidate_genes, this_candidate)
      
    }
  } 
}
colnames(candidate_genes) <- c("SNP", "gene_ID", "functional_annotation", "arabidopsis_annot", "GOterm", "chrom", "startpos", "endpos", "mid_distance")
View( unique(candidate_genes) %>% arrange(SNP, as.numeric(mid_distance)) )

write.csv(unique(candidate_genes) %>% arrange(SNP, as.numeric(mid_distance)), file=paste0(project_path, "output/bivariate_candidate_genes_100kb_revised.csv")) 


sum(candidate_genes$gene_ID %in% apriori_genes$gene_ID)
#no apriori candidates in this list 
```





#### Univariate associations candidate gene analysis: ####



Read in and compile significant SNPs from univariate associations
```{r}
# Set significance cut-off
alpha <- 0.05
#Gao modified Bonferroni
Gao.cutoff <- 1.367e-05

TCICHK_hits <- readRDS(file=paste0(project_path,
    "output/univariate_models/TCICHK_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr)) %>%
  mutate(trait = "TCICHK")

TCHART_hits <- readRDS(file=paste0(project_path,
             "output/univariate_models/TCHART_all_SNP_results_df_revised_weighted_0717.RDS")) %>% 
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "TCHART")

DM_hits <- readRDS(file=paste0(project_path, "output/univariate_models/DM_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "DM")

DYLD_hits <- readRDS(file=paste0(project_path,
             "output/univariate_models/DYLD_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "DYLD")

RTWT_hits <- readRDS(file=paste0(project_path,
             "output/univariate_models/RTWT_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "RTWT")

HI_hits <- readRDS(file=paste0(project_path,
             "output/univariate_models/HI_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "HI")

SPRTPERC_hits <- readRDS(file=paste0(project_path,
             "output/univariate_models/SPRTPERC_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "SPRTPERC")

FRYLD_hits <- readRDS(file=paste0(project_path,
             "output/univariate_models/FRYLD_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "FRYLD")

CMD3S_hits <- readRDS(file=paste0(project_path,
             "output/univariate_models/CMD3S_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "CMD3S")

CMD3I_hits <- readRDS(file=paste0(project_path,
             "output/univariate_models/CMD3I_all_SNP_results_df_revised_weighted_0717.RDS")) %>%
  filter(as.numeric(Pr) <= Gao.cutoff) %>%
  arrange(as.numeric(Pr))%>%
  mutate(trait = "CMD3I")



all_uni_results <- rbind(TCICHK_hits, TCHART_hits, DM_hits, DYLD_hits, RTWT_hits, HI_hits, 
                         SPRTPERC_hits, FRYLD_hits, CMD3S_hits, CMD3I_hits) %>%
  mutate(POS = as.numeric(POS))

```


Identify peak SNP in each QTL
```{r}
hittraitlist <- unique(all_uni_hits$trait)

alltraits_peaks <- data.frame()
for(j in seq_along(hittraitlist)){
  
  traitj <- hittraitlist[j]
  print(traitj)
  traitj_uni_hits <- all_uni_hits %>% filter(trait == traitj)

  chroms_to_seq <- unique(traitj_uni_hits$CHROM)
  
  #go through each chromosome that has sig SNPs
  
  traitj_peaks_allchrom <- data.frame()
  for(i in seq_along(chroms_to_seq)) {
    print(chroms_to_seq[i])
    
    #Identify top SNP on chromosome i 
    topSNPs_chri <- traitj_uni_hits %>% filter(CHROM == chroms_to_seq[i]) %>% arrange(as.numeric(Pr))
    peak1 <- topSNPs_chri[1,]
    print(peak1$SNP_ID)
    
    #check if there is another peak on this chromosome (> 500kb away and not in high LD) 
    peak1_region <- c( peak1$POS - 500000 , peak1$POS + 500000 )  
   
    otherpeaks <- topSNPs_chri %>% filter(POS < peak1_region[1] | POS > peak1_region[2])
    
    if( nrow(otherpeaks) != 0){ #check for LD with peak 1
    
    genomat_peaks_i <- genomat_ROG %>% filter(SNP %in% c(peak1$SNP_ID, otherpeaks$SNP_ID)) 
    rownames(genomat_peaks_i) <- genomat_peaks_i$SNP
    
    LDmatrix_peaks <- genomat_peaks_i %>%
      dplyr::select(-c(SNP, compute_group, CHROM, POS)) %>% #drop metadata non-numeric columns
      t() %>% cor() #calculate correlation matrix
    
    R2withpeak1 <- (LDmatrix_peaks[,paste0(peak1$SNP_ID)])^2 #R2 with peak1 SNP
    
    sigSNPsnotinLDwpeak1 <- R2withpeak1[R2withpeak1 <= 0.05]
    
    peak2 <- otherpeaks %>% filter(SNP_ID %in% names(sigSNPsnotinLDwpeak1)) %>%
      arrange(as.numeric(Pr)) %>% head(n=1)

    if(nrow(peak2) != 0){
    allchromi_peaks  <- rbind(peak1, peak2)
    }else{allchromi_peaks <- peak1} 
    } #end checking for other peaks
    else{ allchromi_peaks <- peak1 }
    
  #join chromosome_i results with the others
  traitj_peaks_allchrom <- rbind(traitj_peaks_allchrom, allchromi_peaks)
    
  } #end chromosome
    
  #join trait_j results with the others
  alltraits_peaks <- rbind(alltraits_peaks, traitj_peaks_allchrom)
  
} #end trait

saveRDS(alltraits_peaks, file="/workdir/ssv42/mvGWAS/output/univariate_models/alltraits_peakSNPs_revised_weighted.RDS")

#Save for supplementary table 4
#get SNP REF/ALT data:
genomat_uni_peakSNPs <- genomat_ROG %>% filter(SNP %in% alltraits_peaks$SNP_ID) 

#re-calculate AF:
# justgenos <- genomat_uni_peakSNPs %>% dplyr::select(-c(SNP, compute_group, CHROM, POS))
# AFfunction <- function(x){sum(x)/(ncol(justgenos)*2)}
# AFs <- apply(justgenos, MARGIN = 1, FUN = AFfunction)
# genomat_uni_peakSNPs$AF <- AFs

#reattach AF, REF/ALT info:
genomat_uni_peakSNPs_stats <- left_join(genomat_uni_peakSNPs, genomat_SNPinfo %>% dplyr::select(-c(CHROM, POS)), 
                                        by=c("SNP"="SNP_ID"))

alltraits_peaks_stats <- left_join(alltraits_peaks, 
                                   genomat_uni_peakSNPs_stats %>% dplyr::select(SNP, REF, ALT, AF), 
                                   by=c("SNP_ID" = "SNP")) 

write.csv(alltraits_peaks_stats, file="/workdir/ssv42/mvGWAS/output/univariate_models/peakSNP_summarystats_revised_weighted_TableS4.csv")

```




Identify candidate genes within specified range of any significantly associated SNPs:
```{r}
#Filter for unique QTL
search_window_bp <- 100000

candidate_genes <- data.frame()
for(i in 1:nrow(alltraits_peaks)){ #iterate across all significant SNPs (not filtered for LD/duplicates)
  print(i)
  chrom_i <- alltraits_peaks$CHROM[i]
  pos_i <- alltraits_peaks$POS[i]
  
  allgenes_nearby <- allgenes %>% filter(chrom == chrom_i &
                                           start_pos > pos_i - search_window_bp &
                                           end_pos < pos_i + search_window_bp )

  if(nrow(allgenes_nearby) != 0){
  candidate_genes_i <- cbind(alltraits_peaks[i,], allgenes_nearby)
  
  candidate_genes <- rbind(candidate_genes, candidate_genes_i)
  }
  
}

#filter for unique gene/trait combinations 
candidate_genes_unique <- candidate_genes %>% unique() %>% group_by(gene_ID, trait) %>%
  slice(which.min(Pr)) 



View(candidate_genes_unique)


# candidate_genes_table <- candidate_genes_unique %>% 
#   left_join((allgenes %>% dplyr::select(gene_ID, function_annot, start_pos) %>% unique()), by=c("gene_annot"= "function_annot", "start_pos"="start_pos")) %>%
#   arrange(SNPneargene) %>%
#   mutate(midpoint = (as.numeric(start_pos) + as.numeric(end_pos))/2) %>% 
#   mutate(distance_midpoint = abs(as.numeric(gsub("S1.*_", "", peak1_SNP)) - midpoint))
# 
# View(candidate_genes_table)
```


Add in distance of candidate genes to peak SNPs
```{r}
all_candidate_genes_peaks <- data.frame()

for(i in 1:nrow(alltraits_peaks)){
  
  peakSNPi <- alltraits_peaks[i, ]
  
  peakSNPi_candgenes <- candidate_genes_unique %>% filter(trait == peakSNPi$trait, CHROM == peakSNPi$CHROM) %>%
    mutate(distance_peakSNPi = case_when( peakSNPi$POS > start_pos & peakSNPi$POS < end_pos ~ 0, #make distance 0 if the SNP is in the gene
                                          #otherwise take abs distance to midpoint of gene
                                          T ~  abs((start_pos + end_pos)/2 - peakSNPi$POS) )) %>%  
    filter(distance_peakSNPi < 100000) %>%
    mutate(peakSNP = peakSNPi$SNP_ID) %>%
    dplyr::select(trait:peakSNP) %>%
    left_join(., peakSNPi)

  
  all_candidate_genes_peaks <- rbind(all_candidate_genes_peaks, peakSNPi_candgenes)
}

View(unique(all_candidate_genes_peaks))

saveRDS(unique(all_candidate_genes_peaks), file="/workdir/ssv42/mvGWAS/output/univariate_models/univariate_candidate_genes_revised_weighted.RDS")

write.csv(unique(all_candidate_genes_peaks), file="/workdir/ssv42/mvGWAS/output/univariate_models/univariate_candidate_genes_revised_weighted.csv")
```













OLD WAY

Flag peaks and calculate LD:
```{r}
#flag peaks and calculate LD 
topSNPs <- all_uni_hits

# genomat_ROG <- readRDS(file=paste0(project_path, "data/genomat_restofgenome_CASS.RDS")) %>%
#   mutate(CHROM= as.integer(gsub("S", "", SNP) %>% gsub("_[0-9].*", "", .)),
#          POS= as.numeric(gsub("^S.*_", "", SNP)))

traitlist <- unique(all_uni_hits$trait)

alltraits_peaks <- data.frame()
genomat_peaks_allchrom <- data.frame()
alltrait_highLD_peakSNPs <- data.frame()

for(j in seq_along(traitlist)){ #iterate through the traits
  print(traitlist[j])
  
  for(i in 1:18) { #iterate through the chromosomes
    print(i)
    genomat_peaks_i <- genomat_ROG %>% filter(CHROM == i)
    rownames(genomat_peaks_i) <- genomat_peaks_i$SNP
    
    topSNPs_chri <- topSNPs %>% filter(CHROM == i) %>% arrange(as.numeric(Pr))
    peak1 <- topSNPs_chri[1,]
  
    #calculate LD between each SNP and the peak 1 SNP
    if(nrow(peak1) != 0 & !is.na(peak1$SNP_ID)){
      LDmatrix_peak1 <- genomat_peaks_i %>% dplyr::select(-c(SNP, ALT, REF, compute_group, CHROM, POS)) %>% #drop metadata non-numeric columns
        t() %>% cor() #calculate correlation matrix
      R2withpeak1 <- (LDmatrix_peak1[,paste0(peak1$SNP_ID)])^2 #pull out correlation with peak1 SNP
      genomat_peaks_i$R2peak1 <- R2withpeak1
      
      genomat_peaks_i <- genomat_peaks_i %>% mutate(distance_peak1 = POS - peak1$POS, peak1_SNP = peak1$SNP_ID)
    } else{genomat_peaks_i <- genomat_peaks_i %>% mutate(R2peak1 = NA, distance_peak1 = NA, peak1_SNP = NA)}
  
    #check if there is another peak on this chromosome and repeat if so
    peak1_region <- c( peak1$POS - 2000000 , peak1$POS + 2000000)  
    genomat_around_peak1 <- genomat_peaks_i %>% 
         filter(POS > peak1_region[1] & POS < peak1_region[2])
    
    otherpeaks <- topSNPs_chri[!topSNPs_chri$SNP_ID %in% genomat_around_peak1$SNP,]
    peak2 <- otherpeaks %>% arrange(as.numeric(Pr)) %>% head(n=1)
    
    #calculate LD between each SNP and the peak 2 SNP
    if(nrow(peak2) != 0){
      LDmatrix_peak2 <- genomat_peaks_i %>% dplyr::select(-c(SNP, ALT, REF, compute_group, CHROM, POS, distance_peak1, peak1_SNP)) %>% 
        t() %>% cor()
      R2withpeak2 <- (LDmatrix_peak2[,paste0(peak2$SNP_ID)])^2
      genomat_peaks_i$R2peak2 <- R2withpeak2
      
      genomat_peaks_i <- genomat_peaks_i %>% mutate(distance_peak2 = POS - peak2$POS, peak2_SNP = peak2$SNP_ID)
    } else{genomat_peaks_i <- genomat_peaks_i %>% mutate(R2peak2 = NA, distance_peak2 = NA, peak2_SNP = NA)
    }
    
    #join chromosome results with the others
   genomat_peaks_allchrom <- rbind(genomat_peaks_allchrom, genomat_peaks_i)
   
  } #end chromosome
  
  traitj_highLD_SNPs <- genomat_peaks_allchrom %>% filter(R2peak1 > 0.75 | R2peak2 > 0.75) %>%
    mutate(trait = traitlist[j])
  
  alltrait_highLD_peakSNPs <- rbind(alltrait_highLD_peakSNPs, traitj_highLD_SNPs)
  
}#end trait


```



# Get set of candidate genes near top SNPs from univariate associations:
```{r}
#look near SNPs in high LD (> 0.75 of peak) and within 500kb of peak SNP
alltrait_highLD_near_peakSNPs <- alltrait_highLD_peakSNPs %>% 
  filter(abs(distance_peak1 | distance_peak2) < 500000) 

candidate_genes <- data.frame()
for(i in 1:nrow(alltrait_highLD_near_peakSNPs)){ 
  
  #for each SNP near  genes
  
  chrom_i <- alltrait_highLD_near_peakSNPs$CHROM[i]
  pos_i <- alltrait_highLD_near_peakSNPs$POS[i]
  
  allgenes_chrom_i <- allgenes %>% filter(chrom == chrom_i)
  for(j in 1:nrow(allgenes_chrom_i)){
    
    start_search <- allgenes_chrom_i$start_pos[j] - 5000
    end_search <- allgenes_chrom_i$end_pos[j] + 5000
    
    if( pos_i >= start_search && pos_i <= end_search){
    
      this_candidate <- c(as.character(alltrait_highLD_near_peakSNPs$SNP[i]),
                          allgenes_chrom_i$function_annot[j],
                          allgenes_chrom_i$chrom[j],
                          allgenes_chrom_i$start_pos[j],
                          allgenes_chrom_i$end_pos[j],
                          
                          alltrait_highLD_near_peakSNPs$trait[i],
                          
                   alltrait_highLD_near_peakSNPs$peak1_SNP[i],
                   alltrait_highLD_near_peakSNPs$R2peak1[i],
                   alltrait_highLD_near_peakSNPs$distance_peak1[i],
                   
                   alltrait_highLD_near_peakSNPs$peak2_SNP[i],
                   alltrait_highLD_near_peakSNPs$R2peak2[i],
                   alltrait_highLD_near_peakSNPs$distance_peak2[i])
      
      candidate_genes <- rbind(candidate_genes, this_candidate)
      
    }
  } 
}

colnames(candidate_genes) <- c("SNPneargene", "gene_annot", "chromosome", "startpos", "endpos", "trait",
                               "peak1_SNP", "R2peak1", "distance_peak1", 
                               "peak2_SNP", "R2peak2", "distance_peak2")

candidate_genes_unique <- candidate_genes %>% unique() %>% group_by(gene_annot) %>%
  slice(which.max(max(c(R2peak1, R2peak2))))  %>%
  mutate(start_pos = as.numeric(startpos), end_pos=as.numeric(endpos))

candidate_genes_table <- candidate_genes_unique %>% 
  left_join((allgenes %>% dplyr::select(gene_ID, function_annot, start_pos) %>% unique()), by=c("gene_annot"= "function_annot", "start_pos"="start_pos")) %>%
  arrange(SNPneargene) %>%
  mutate(midpoint = (as.numeric(start_pos) + as.numeric(end_pos))/2) %>% 
  mutate(distance_midpoint = abs(as.numeric(gsub("S1.*_", "", peak1_SNP)) - midpoint))

View(candidate_genes_table)

```


Identify candidate genes from univariate associations:
```{r}
#Filter for unique QTL

search_window_bp <- 100000

candidate_genes <- data.frame()
for(i in 1:nrow(alltraits_peaks)){ #iterate across all significant SNPs (not filtered for LD/duplicates yet)
  print(i)
  chrom_i <- alltraits_peaks$CHROM[i]
  pos_i <- as.numeric(alltraits_peaks$POS[i])
  
  allgenes_nearby <- allgenes %>% filter(chrom == chrom_i &
                                           start_pos > pos_i - search_window_bp &
                                           end_pos < pos_i + search_window_bp )
  if(nrow(allgenes_nearby != 0)){
  candidate_genes_i <- cbind(alltraits_peaks[i,], allgenes_nearby)
  
  candidate_genes <- rbind(candidate_genes, candidate_genes_i)
  }
  
}

#filter for unique gene/trait combinations 

candidate_genes_unique <- candidate_genes %>% unique() %>% group_by(gene_ID, trait) %>%
  slice(which.min(Pr)) 

View(candidate_genes_unique)


# candidate_genes_table <- candidate_genes_unique %>% 
#   left_join((allgenes %>% dplyr::select(gene_ID, function_annot, start_pos) %>% unique()), by=c("gene_annot"= "function_annot", "start_pos"="start_pos")) %>%
#   arrange(SNPneargene) %>%
#   mutate(midpoint = (as.numeric(start_pos) + as.numeric(end_pos))/2) %>% 
#   mutate(distance_midpoint = abs(as.numeric(gsub("S1.*_", "", peak1_SNP)) - midpoint))
# 
# View(candidate_genes_table)
```









Old way:






